### 2023-12-15 tex/abstract.tex
commit some tex stuff

%!TEX root = main.tex

We need to write an abstract.