### 2023-12-15 tex/author-title-etc.tex

%!TEX root = main.tex

%%%%% THIS IS THE SECTION WHERE THE AUTHOR PUTS IN ALL OF THEIR TITLE AND AFFILIATION
%%%%% INFORMATION AND  A FEW OTHER THINGS
\newcommand{\myTitle}{A multipurpose microhaplotype panel for analysis of California Chinook salmon}
\title{\myTitle}

% redefine this to make author list. Note affiliation symbols are done manually.
% All caps tends to look better
\newcommand{\myAuthors}{Eric C. Anderson$^{*,\S}$, Anthony J. Clemento$^{*}$, Matthew A. Campell$^{?}$, Devon Pearse$^{*}$, Cassie Columbus$^{*}$, Ellen Campbell$^{*}$, John Carlos Garza$^{*}$}
\author{Eric C. Anderson$^{*}$, Anthony J. Clemento$^{*}$, Matthew A. Campell$^{?}$, Devon Pearse$^{*}$, Cassie Columbus$^{*}$, Ellen Campbell$^{*}$, John Carlos Garza$^{*}$}

% redefine to make the affiliation list.  Note symbols are done manually
\newcommand{\myAffiliations}{
$^*$Southwest Fisheries Science Center, National Marine Fisheries Service, NOAA,Santa Cruz, California, USA.
}

\renewcommand{\AuthorAddresses}{\myAffiliations}

\renewcommand{\KeyWords}{Genetic stock identfication, population assignment, parentage based tagging, amplicon sequencing}

\renewcommand{\CorrespondingAuthor}{eric.c.anderson@noaa.gov}

% the email address for the corresponding author
\newcommand{\myEmailAddress}{eric.c.anderson@noaa.gov}
\newcommand{\myEmailFootnote}{$^\S$}

% here you can put your very own copyright notice
\newcommand{\myCopyright}{\copyright US Federal Government work in the public domain in the USA}

% here you can put a running title (a short title that goes on the left of the
% even pages)
\newcommand{\myRunningTitle}{Microhaplotypes for California Chinook salmon}
\renewcommand{\RunningTitle}{\myRunningTitle}

% and here you put the running author (short listing of authors for the
% upper right header on the odd pages)
\newcommand{\myRunningAuthor}{Anderson et al.}

%%%% DONE WITH AUTHOR/TITLE/ETC INFORMATION DEFINITION