### 2023-12-15 tex/data-accessibility.tex

\section*{Data Accessibility}

{\sc WGSassign} is available as a Python package with these associated links:
\begin{itemize}
\item Development version and entire revision history on GitHub: \url{https://anonymous.4open.science/r/WGSassign-7155}
\item Zenodo archive of initial package release: UPDATE AFTER ANONYMOUS REVIEW
\item Online version of data and scripts used in paper: \url{https://anonymous.4open.science/r/WGSassign-manuscript-data-3072}
\item Data repository with full datasets used in paper. UPDATED WHEN JOURNAL PROVIDES DOI. \url{https://dryad.something.or.other}
\end{itemize}
\mbox{}