### 2023-12-20 map-of-samples.Rmd
Working on the map

---
title: "Make a Map for our Paper"
author: "Eric C. Anderson"
date: "Last Updated: `r Sys.Date()`"
output:
  html_document:
    df_print: paged
    toc: true
    toc_float: true
bibliography: "`r here::here('references.bib')`"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(message = FALSE)
knitr::opts_knit$set(root.dir = here::here())
start_time <- Sys.time()

USING_SNAKEMAKE <- FALSE
NEG_USING_SMK <- TRUE
if(exists("snakemake")) {
  USING_SNAKEMAKE <- TRUE
  NEG_USING_SMK <- FALSE
}
```

```{r rmd_snakemake_hacks, include=FALSE, eval=USING_SNAKEMAKE}
# currently, this block does not get evaluated, while developing,
# but later when we snakemake-ize it all, we will evaluate it.
if(USING_SNAKEMAKE) {
  input_list <- snakemake@input
  output_list <- snakemake@output
}
```

# Input and Output Paths

```{r}
# inputs:
input_list <- list(
)
# outputs:
output_list <- list(
)
# you can create the necessary output directories like this:
dump <- lapply(output_list, function(x)
  dir.create(dirname(x), recursive = TRUE, showWarnings = FALSE)
)
```

```{r}
library(terra)
library(sf)
library(tidyverse)
library(ggspatial)
#library(ggsn)
library(cowplot)
library(plotly)
```

# Download spatial data

## Raster

```{r}
if(!file.exists("geo-spatial/HYP_HR_SR_OB_DR/HYP_HR_SR_OB_DR.tif")) {
  dir.create("geo-spatial", showWarnings = FALSE)
  download.file(
    url = "https://www.naturalearthdata.com/http//www.naturalearthdata.com/download/10m/raster/HYP_HR_SR_OB_DR.zip",
    destfile = "geo-spatial/HYP_HR_SR_OB_DR.zip"
  )

  unzip(
    zipfile = "geo-spatial/HYP_HR_SR_OB_DR.zip", 
    exdir = "geo-spatial/HYP_HR_SR_OB_DR"
  )

  file.remove("geo-spatial/HYP_HR_SR_OB_DR.zip")
}
```

## State lines

```{r}
if(!file.exists("geo-spatial/ne_10m_admin_1_states_provinces_lines/ne_10m_admin_1_states_provinces_lines.shp")) {
  dir.create("geo-spatial", showWarnings = FALSE)
  download.file(
    url = "https://www.naturalearthdata.com/http//www.naturalearthdata.com/download/10m/cultural/ne_10m_admin_1_states_provinces_lines.zip",
    destfile = "geo-spatial/ne_10m_admin_1_states_provinces_lines.zip"
  )

  unzip(
    zipfile = "geo-spatial/ne_10m_admin_1_states_provinces_lines.zip", 
    exdir = "geo-spatial/ne_10m_admin_1_states_provinces_lines"
  )

  file.remove("geo-spatial/ne_10m_admin_1_states_provinces_lines.zip")
}
```

## Coastline

```{r}
if(!file.exists("geo-spatial/ne_10m_coastline/ne_10m_coastline.shp")) {
  dir.create("geo-spatial", showWarnings = FALSE)
  download.file(
    url = "https://www.naturalearthdata.com/http//www.naturalearthdata.com/download/10m/physical/ne_10m_coastline.zip",
    destfile = "geo-spatial/ne_10m_coastline.zip"
  )

  unzip(
    zipfile = "geo-spatial/ne_10m_coastline.zip", 
    exdir = "geo-spatial/ne_10m_coastline"
  )

  file.remove("geo-spatial/ne_10m_coastline.zip")
}
```

## California Major Rivers and Creeks

This is from California Nat Res Agency.  Find it at: https://data.cnra.ca.gov/dataset/national-hydrography-dataset-nhd

```{r}
# note.  Once when I did this, download.file did not see capable of downloading this file on my Mac
# (the download is super sloooooow), and that time I 
# ended up downloading it with Chrome.  But it seems those connection problems
# had been fixed when I did this again.
if(!file.exists("geo-spatial/NHD_Major_Rivers_and_Creeks/Major_Rivers_and_Creeks.shp")) {
  dir.create("geo-spatial", showWarnings = FALSE)
  download.file(
    url = "https://data.cnra.ca.gov/dataset/511528b2-f7d3-4d86-8902-cc9befeeeed5/resource/7d1e7e44-81b1-43fe-95f6-1862eea6ac24/download/nhd_major_rivers_and_creeks.zip",
    destfile = "geo-spatial/nhd_major_rivers_and_creeks.zip"
  )

  unzip(
    zipfile = "geo-spatial/nhd_major_rivers_and_creeks.zip", 
    exdir = "geo-spatial"
  )

  file.remove("geo-spatial/nhd_major_rivers_and_creeks.zip")
}
```

# Read in the spatial data

```{r}
nat.earth <- rast("geo-spatial/HYP_HR_SR_OB_DR/HYP_HR_SR_OB_DR.tif") #
state_prov <- st_read("geo-spatial/ne_10m_admin_1_states_provinces_lines/ne_10m_admin_1_states_provinces_lines.shp")
coastline <- st_read("geo-spatial/ne_10m_coastline/ne_10m_coastline.shp")
all_rivers <- st_read("geo-spatial/NHD_Major_Rivers_and_Creeks/Major_Rivers_and_Creeks.shp") %>%
  st_zm() %>%
  st_transform(., st_crs(state_prov))
```

## Subset the rivers

We plot this with plotly so we can see the names of the rivers easily, and whittle
it down to the ones that we need.

```{r}
# I don't know the gnis_id for mill creek, and there are a lot of Mill Creeks,
# so lets get it:
mill_area <- c(
  xmin = -121.4,
  xmax = -122.3,
  ymin = 40,
  ymax = 40.3
)
mill_candi <- all_rivers %>%
  filter(str_detect(gnis_name, "Mill Creek")) %>%
  st_crop(mill_area)

our_rivers <- all_rivers %>%
  filter(gnis_name %in% c(
    "Sacramento River",
    "San Joaquin River",
    "Feather River",
    "Russian River",
    "Trinity River"
    #"Deer Creek",
    #"Battle Creek"
  ) | 
    gnis_id %in% c(
      "00220293", "00237425", "00263498", "00266522",  # these are butte creek
      "00233775", "01655075",    # these are deer
      "00218740", "00229640", "00234966"   # these are battle creek
    ) 
  ) %>%
  bind_rows(mill_candi)

# get candidate positions for notations
labels <- read_tsv("inputs/map-notations.tsv")

g <- ggplot() + 
  geom_sf(data = our_rivers, aes(colour = gnis_name)) +
  geom_point(data = labels, aes(x = label_long, y = label_lat), colour = "red") +
  geom_point(data = labels, aes(x = tip_long, y = tip_lat), colour = "blue", size = 0.2)

ggplotly(g)
```

That looks good.  So let us proceed.

## Crop
```{r}
# important to put them in this order and named like this
domain <- c(
  xmin = -128,
  xmax = -116,
  ymin = 36,
  ymax = 42.5
)
nat_crop <- crop(nat.earth, y = ext(domain))
state_subset <- st_crop(state_prov, domain)
coastline_cropped <- st_crop(coastline, domain)

```

# Make the base map
```{r base-map-with-sampling-sites}
base_map <- ggplot() +
  ggspatial::layer_spatial(nat_crop) +
  geom_sf(data = state_subset, color = "gray30", fill = NA) +
  geom_sf(data = coastline_cropped, color = "gray30", fill = NA) +
  geom_sf(data = our_rivers, colour = "blue")

base_map
```

Now look add stuff to that:

```{r}
base_map + 
  geom_point(data = labels, aes(x = label_long, y = label_lat), colour = "red") +
  geom_point(data = labels, aes(x = tip_long, y = tip_lat), colour = "blue", size = 0.2)
```

# Grab data for sampling sites

```{r}
fish_sites <- read_csv("./data/100-RoSA_sample_sites.csv")
```

```{r}

  geom_text(data = fish_sites, aes(x = name_long, y = name_lat, label = pop, family = "serif", fontface = font_type), size = 3.5) +
  geom_segment(data = fish_sites %>% filter(!pop %in% c("San Joaquin", "River (F)")), mapping = aes(x = line_long, xend = longitude, y = name_lat - 0.05, yend = latitude), size = 0.5) +
  geom_segment(data = fish_sites %>% filter(pop == "San Joaquin"), mapping = aes(x = line_long, xend = longitude, y = name_lat - 0.55, yend = latitude), size = 0.5) +
  geom_point(data = fish_sites, mapping = aes(x = longitude, y = latitude), fill = fish_sites$color2, colour = fish_sites$color2) +
  scale_x_continuous(expand = c(0, 0)) +
  xlab("Longitude") +
  scale_y_continuous("Latitude", expand = c(0, 0)) +
  coord_sf(xlim = domain[1:2], ylim = domain[3:4]) +
  scalebar(
    x.min = domain[1], x.max = domain[2], y.min = domain[3], y.max = domain[4],
    location = "bottomright", model = "WGS84", dist = 250, anchor = c(x = -123, y = 36), st.size = 2.5, transform = TRUE, dist_unit = "km"
  ) +
  theme_bw() +
  theme(
    panel.border = element_rect(colour = "black", size = 1),
    axis.text.x = element_text(size = 8, family = "serif", angle = 35, hjust = 1),
    axis.text.y = element_text(size = 8, family = "serif"),
    axis.title.y = element_text(family = "serif", size = 10),
    axis.title.x = element_text(family = "serif", vjust = 2, size = 10),
    plot.margin = margin(0, 0.1, 0, 0.15, "cm"),
    legend.position = "none"
  )

# base_map
```

## add north arrow to map.
```{r adding-arrow-to-base-map}
arrow_map <- base_map +
  north(x.min = domain[-1], x.max = domain[2], y.min = domain[3], y.max = domain[4], location = "bottomright", anchor = c(x = -120.25, y = 36.5))

# arrow_map
```

## Now, work on the world-scale map with the inset:
```{r inset-map}
wrld <- map_data("state")
domain_df <- data_frame(point = 1:length(domain), long = rep(domain[1:2], each = 2), lat = c(domain[3:4], rev(domain[3:4])))
inset_world <- ggplot() +
  geom_path(data = wrld, aes(x = long, y = lat, group = group), colour = "black", size = 0.1) +
  geom_polygon(data = domain_df, mapping = aes(x = long, y = lat), colour = "red", fill = "red", alpha = 0.3) +
  coord_map("ortho", orientation = c(41, -132, 0)) +
  theme_bw() +
  labs(x = NULL, y = NULL) +
  theme(
    axis.title = element_blank(),
    axis.text = element_blank(),
    axis.ticks = element_blank(),
    plot.margin = unit(c(0, 0, -1, -1), "mm")
  )
# inset_world
```

## Use cowplot to make the inset appear on the bigger map

```{r cowplot-to-make-inset}
final_map <- ggdraw() +
  draw_plot(arrow_map) +
  draw_plot(inset_world, x = 0.7, y = 0.725, width = 0.25, height = 0.2)

ggsave(final_map, filename = "./outputs/100/RoSA_figure1_map_with_inset.pdf", width = 4.75, height = 4.25)
# and then plot it in the notebook (albeit at different size) too
final_map
```

# Citations

::: {#refs}
:::

# Session Info

```{r}
sessioninfo::session_info()
```

# Running Time

Running the code and rendering this notebook required approximately this much time:

```{r}
Sys.time() - start_time
```