### 2023-12-21 tex/newcomms.tex

\newcommand{\relat}{R}
\newcommand{\relatset}{\mathscr{R}}
\newcommand{\nA}{\mathcal{A}}
\newcommand{\nL}{\mathcal{L}}
\newcommand{\nG}{\mathcal{G}}
\newcommand{\MAF}{\mathrm{MAF}}

\newcommand{\Pe}{\mathrm{P}_\epsilon}
\newcommand{\PR}{\mathrm{P}_\relat}

\newcommand{\bkappa}{\bm{\kappa}}

\newcommand{\comment}[1]{\textcolor{red}{#1}}

\newcommand{\BY}{\mathbf{Y}}
\newcommand{\BX}{\mathbf{X}}
\newcommand{\BC}{\mathbf{C}}
\newcommand{\I}{\mathbb{1}}
\newcommand{\bomega}{\bm{\omega}}

\newcommand{\Exp}{\mathbb{E}}
\newcommand{\Var}{\mathrm{Var}}

\newcommand{\heta}{\hat{\theta}}
\newcommand{\bg}{\bm{g}}

\newcommand{\GG}{\mathscr{G}}

\newcommand{\LFball}{\raisebox{-0.12 em}{\includegraphics[height=1.02em]{images/Late-fall-run-ball-crop.pdf}}}
\newcommand{\Fball}{\raisebox{-0.12 em}{\includegraphics[height=1.02em]{images/Fall-run-ball-crop.pdf}}}
\newcommand{\Wball}{\raisebox{-0.12 em}{\includegraphics[height=1.02em]{images/Winter-run-ball-crop.pdf}}}
\newcommand{\Sball}{\raisebox{-0.12 em}{\includegraphics[height=1.02em]{images/Spring-run-ball-crop.pdf}}}