### 2024-01-07 map-of-samples.Rmd

---
title: "Make a Map for our Paper"
author: "Eric C. Anderson"
date: "Last Updated: `r Sys.Date()`"
output:
  html_document:
    df_print: paged
    toc: true
    toc_float: true
bibliography: "`r here::here('tex/citation.bib')`"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(message = FALSE)
knitr::opts_knit$set(root.dir = here::here())
start_time <- Sys.time()

USING_SNAKEMAKE <- FALSE
NEG_USING_SMK <- TRUE
if(exists("snakemake")) {
  USING_SNAKEMAKE <- TRUE
  NEG_USING_SMK <- FALSE
}
```

```{r rmd_snakemake_hacks, include=FALSE, eval=USING_SNAKEMAKE}
# currently, this block does not get evaluated, while developing,
# but later when we snakemake-ize it all, we will evaluate it.
if(USING_SNAKEMAKE) {
  input_list <- snakemake@input
  output_list <- snakemake@output
}
```

# Input and Output Paths

```{r}
# inputs:
input_list <- list(
)
# outputs:
output_list <- list(
  map = "results/map_of_samples/map.pdf",
  cropmap = "results/map_of_samples/map-crop.pdf",
  texmap = "tex/images/map-crop.pdf"
)
# you can create the necessary output directories like this:
dump <- lapply(output_list, function(x)
  dir.create(dirname(x), recursive = TRUE, showWarnings = FALSE)
)
```

```{r}
library(terra)
library(sf)
library(tidyverse)
library(ggspatial)
#library(ggsn)
library(cowplot)
library(plotly)
```

# Download spatial data

## Raster

```{r}
if(!file.exists("geo-spatial/HYP_HR_SR_OB_DR/HYP_HR_SR_OB_DR.tif")) {
  dir.create("geo-spatial", showWarnings = FALSE)
  download.file(
    url = "https://www.naturalearthdata.com/http//www.naturalearthdata.com/download/10m/raster/HYP_HR_SR_OB_DR.zip",
    destfile = "geo-spatial/HYP_HR_SR_OB_DR.zip"
  )

  unzip(
    zipfile = "geo-spatial/HYP_HR_SR_OB_DR.zip", 
    exdir = "geo-spatial/HYP_HR_SR_OB_DR"
  )

  file.remove("geo-spatial/HYP_HR_SR_OB_DR.zip")
}
```

## State lines

```{r}
if(!file.exists("geo-spatial/ne_10m_admin_1_states_provinces_lines/ne_10m_admin_1_states_provinces_lines.shp")) {
  dir.create("geo-spatial", showWarnings = FALSE)
  download.file(
    url = "https://www.naturalearthdata.com/http//www.naturalearthdata.com/download/10m/cultural/ne_10m_admin_1_states_provinces_lines.zip",
    destfile = "geo-spatial/ne_10m_admin_1_states_provinces_lines.zip"
  )

  unzip(
    zipfile = "geo-spatial/ne_10m_admin_1_states_provinces_lines.zip", 
    exdir = "geo-spatial/ne_10m_admin_1_states_provinces_lines"
  )

  file.remove("geo-spatial/ne_10m_admin_1_states_provinces_lines.zip")
}
```

## Coastline

```{r}
if(!file.exists("geo-spatial/ne_10m_coastline/ne_10m_coastline.shp")) {
  dir.create("geo-spatial", showWarnings = FALSE)
  download.file(
    url = "https://www.naturalearthdata.com/http//www.naturalearthdata.com/download/10m/physical/ne_10m_coastline.zip",
    destfile = "geo-spatial/ne_10m_coastline.zip"
  )

  unzip(
    zipfile = "geo-spatial/ne_10m_coastline.zip", 
    exdir = "geo-spatial/ne_10m_coastline"
  )

  file.remove("geo-spatial/ne_10m_coastline.zip")
}
```

## California Major Rivers and Creeks

This is from California Nat Res Agency.  Find it at: https://data.cnra.ca.gov/dataset/national-hydrography-dataset-nhd

```{r}
# note.  Once when I did this, download.file did not see capable of downloading this file on my Mac
# (the download is super sloooooow), and that time I 
# ended up downloading it with Chrome.  But it seems those connection problems
# had been fixed when I did this again.
if(!file.exists("geo-spatial/NHD_Major_Rivers_and_Creeks/Major_Rivers_and_Creeks.shp")) {
  dir.create("geo-spatial", showWarnings = FALSE)
  download.file(
    url = "https://data.cnra.ca.gov/dataset/511528b2-f7d3-4d86-8902-cc9befeeeed5/resource/7d1e7e44-81b1-43fe-95f6-1862eea6ac24/download/nhd_major_rivers_and_creeks.zip",
    destfile = "geo-spatial/nhd_major_rivers_and_creeks.zip"
  )

  unzip(
    zipfile = "geo-spatial/nhd_major_rivers_and_creeks.zip", 
    exdir = "geo-spatial"
  )

  file.remove("geo-spatial/nhd_major_rivers_and_creeks.zip")
}
```

# Read in the spatial data

```{r}
nat.earth <- rast("geo-spatial/HYP_HR_SR_OB_DR/HYP_HR_SR_OB_DR.tif") #
state_prov <- st_read("geo-spatial/ne_10m_admin_1_states_provinces_lines/ne_10m_admin_1_states_provinces_lines.shp")
coastline <- st_read("geo-spatial/ne_10m_coastline/ne_10m_coastline.shp")
all_rivers <- st_read("geo-spatial/NHD_Major_Rivers_and_Creeks/Major_Rivers_and_Creeks.shp") %>%
  st_zm() %>%
  st_transform(., st_crs(state_prov))
```

## Subset the rivers

We plot this with plotly so we can see the names of the rivers easily, and whittle
it down to the ones that we need.

```{r}
# I don't know the gnis_id for mill creek, and there are a lot of Mill Creeks,
# so lets get it:
mill_area <- c(
  xmin = -121.4,
  xmax = -122.3,
  ymin = 40,
  ymax = 40.3
)
mill_candi <- all_rivers %>%
  filter(str_detect(gnis_name, "Mill Creek")) %>%
  st_crop(mill_area)

our_rivers <- all_rivers %>%
  filter(gnis_name %in% c(
    "Sacramento River",
    "San Joaquin River",
    "Feather River",
    "Russian River",
    "Trinity River",
    "Klamath River"
    #"Deer Creek",
    #"Battle Creek"
  ) | 
    gnis_id %in% c(
      "00220293", "00237425", "00263498", "00266522",  # these are butte creek
      "00233775", "01655075"    # these are deer
     # "00218740", "00229640", "00234966"   # these are battle creek
    ) 
  ) %>%
  bind_rows(mill_candi)

# get candidate positions for notations
labels <- read_tsv("inputs/map-notations.tsv", comment = "#")

g <- ggplot() + 
  geom_sf(data = our_rivers, aes(colour = gnis_name)) +
  geom_point(data = labels, aes(x = label_long, y = label_lat), colour = "red") +
  geom_point(data = labels, aes(x = tip_long, y = tip_lat), colour = "blue", size = 0.2)

ggplotly(g)
```

That looks good.  So let us proceed.

## Crop
```{r}
# important to put them in this order and named like this
domain <- c(
  xmin = -127.9,
  xmax = -117.4,
  ymin = 36,
  ymax = 42.6
)
sf_use_s2(FALSE)
nat_crop <- crop(nat.earth, y = ext(domain))
state_subset <- st_crop(state_prov, domain)
coastline_cropped <- st_crop(coastline, domain)

```

# Make the base map
```{r base-map-with-sampling-sites}
base_map <- ggplot() +
  ggspatial::layer_spatial(nat_crop) +
  geom_sf(data = state_subset, color = "gray30", fill = NA) +
  geom_sf(data = coastline_cropped, color = "gray30", fill = NA, linewidth = 0.15) +
  geom_sf(data = our_rivers, colour = "blue", linejoin = "round", lineend = "round")

```

Now look add stuff to that:

```{r}
bit <- 0.0
kick <- 0.4
source("R/colors.R")

labels_1line <- labels %>%
  group_by(map_text) %>%
  filter(rank == max(rank)) %>%
  ungroup()

mm <- base_map + 
  geom_segment(data = labels, aes(x = tip_long, y = tip_lat, xend = label_long, yend = label_lat), colour = "black", size = 0.4) +
  geom_point(data = labels, aes(x = label_long + bit + (1 - hjust * 2) * kick * (rank - 1), y = label_lat, fill = run_timing), shape = 21, size = 3.5) +
  geom_label(
    data = labels_1line,
    aes(
      x = label_long + bit + (1 - hjust * 2) * kick * ( 0.65 + rank - 1),
      y = label_lat, 
      label = map_text,
      hjust = hjust
    ),
    size = 2.2,
    label.padding = unit(0.09, "lines"),
    lineheight = 0.85
  ) +
  scale_fill_manual(values = run_time_colors) +
  theme_bw() +
  theme(
    panel.border = element_rect(colour = "black", linewidth = 1),
    axis.text.x = element_text(size = 8, family = "serif", angle = 35, hjust = 1),
    axis.text.y = element_text(size = 8, family = "serif"),
    axis.title.y = element_text(family = "serif", size = 10),
    axis.title.x = element_text(family = "serif", vjust = 2, size = 10),
    plot.margin = margin(0, 0.1, 0, 0.15, "cm"),
    legend.position = "none"
  ) +
  xlab("Longitude") +
  ylab("Latitude") +
  coord_sf(
    expand = FALSE,
  ) +
  guides(fill = guide_legend(title = "Run Timing:", nrow = 2)) +
  ggspatial::annotation_north_arrow(location = "br", height = unit(10, "mm"), width = unit(5, "mm"), style = north_arrow_fancy_orienteering()) +
  ggspatial::annotation_scale(location = "br", height = unit(0.1, "cm"))

```

## Now, work on the world-scale map with the inset:
```{r inset-map}
wrld <- map_data("world")
states <- map_data("state")
domain_df <- tibble(point = 1:length(domain), long = rep(domain[1:2], each = 2), lat = c(domain[3:4], rev(domain[3:4])))

inset_world <- ggplot() +
  geom_polygon(data = wrld, aes(x = long, y = lat, group = group), colour = "black", fill = "gray90", size = 0.1) +
  geom_path(data = states, aes(x = long, y = lat, group = group), colour = "black", size = 0.1) +
  geom_polygon(data = domain_df, mapping = aes(x = long, y = lat), colour = "red", fill = "red", alpha = 0.3, linewidth = 0.1) +
  coord_map("ortho", orientation = c(30, -114, 0), xlim = c(-150, -75), ylim = c(28, 60)) +
  theme_bw() +
  labs(x = NULL, y = NULL) +
  theme(
    axis.title = element_blank(),
    axis.text = element_blank(),
    axis.ticks = element_blank(),
    plot.margin = unit(c(0, 0, -1, -1), "mm")
  )
inset_world
```

## Use cowplot to make the inset appear on the bigger map

```{r cowplot-to-make-inset}
final_map <- ggdraw() +
  draw_plot(mm) +
  draw_plot(inset_world, x = 0.181, y = 0.139, width = 0.20, height = 0.17)

ggsave(final_map, filename = output_list$map, width = 5, height = 3.5)

# let's crop that down for use, too
CALL <- paste("pdfcrop ", output_list$map, collapse = " ")
system(CALL)

file.rename(from = output_list$cropmap, to = output_list$texmap)

```

# Make colored balls for inclusion into the caption

We can do this and just kick them directly into the tex/images folder.  
```{r}
rc_col_tib <- enframe(run_time_colors) %>%
  mutate(
    nospace = str_replace_all(name, " ", "-"),
    file = str_c("tex/images/", nospace, "-ball.pdf")
  )

for(r in 1:nrow(rc_col_tib)) {
  tmp <- rc_col_tib[r,]
  g <- ggplot(tmp, aes(x = 1, y = 1)) +
    geom_point(shape = 21, fill = tmp$value, size = 3.5) +
    theme_void()
  ggsave(g, filename = tmp$file)

  CALL <- str_c("pdfcrop ", tmp$file)
  system(CALL)
  file.remove(tmp$file)
}
```

# Citations

::: {#refs}
:::

# Session Info

```{r}
sessioninfo::session_info()
```

# Running Time

Running the code and rendering this notebook required approximately this much time:

```{r}
Sys.time() - start_time
```