### 2024-01-07 tex/inputs/samples-table.tex

Mainstem River$^1$&Code&Location Name$^2$&Run&ESU$^3$&$N$&Sampling Months&Year Range\tabularnewline
\hline Sacramento R.&BCF&Butte Ck.&Fall&CVFR&45&\raisebox{-0.12 em}{\includegraphics[height=1.02em]{../tex/images/months-BCF_Butte-Creek.pdf}}&2004--2016\tabularnewline
Sacramento R.&BCS&Butte Ck.&Spring&CVSR&116&\raisebox{-0.12 em}{\includegraphics[height=1.02em]{../tex/images/months-BCS_Butte-Creek.pdf}}&2004--2017\tabularnewline
Sacramento R.&CHLF&Coleman NFH&Late-fall&CVFR&261&\raisebox{-0.12 em}{\includegraphics[height=1.02em]{../tex/images/months-CHLF_Coleman-National-Fish-Hatchery.pdf}}&2011--2015\tabularnewline
Sacramento R.&MDF&Deer Ck.&Fall&CVFR&29&\raisebox{-0.12 em}{\includegraphics[height=1.02em]{../tex/images/months-MDF_Deer-Creek.pdf}}&2002--2004\tabularnewline
Sacramento R.&MDS&Deer Ck.&Spring&CVSR&34&\raisebox{-0.12 em}{\includegraphics[height=1.02em]{../tex/images/months-MDS_Deer-Creek.pdf}}&2002--2002\tabularnewline
Sacramento R.&FRHF&Feather R. H.&Fall&CVFR&106&\raisebox{-0.12 em}{\includegraphics[height=1.02em]{../tex/images/months-FRHF_Feather-River-Hatchery.pdf}}&2008--2010\tabularnewline
Sacramento R.&FRHS&Feather R. H.&Spring&CVSR&202&\raisebox{-0.12 em}{\includegraphics[height=1.02em]{../tex/images/months-FRHS_Feather-River-Hatchery.pdf}}&2010--2019\tabularnewline
Sacramento R.&MDF&Mill Ck.&Fall&CVFR&53&\raisebox{-0.12 em}{\includegraphics[height=1.02em]{../tex/images/months-MDF_Mill-Creek.pdf}}&2002--2004\tabularnewline
Sacramento R.&MDS&Mill Ck.&Spring&CVSR&51&\raisebox{-0.12 em}{\includegraphics[height=1.02em]{../tex/images/months-MDS_Mill-Creek.pdf}}&2002--2004\tabularnewline
Russian R.&RRF&Warm Springs H.&Fall&CC&92&\raisebox{-0.12 em}{\includegraphics[height=1.02em]{../tex/images/months-RRF_Warm-Springs-Hatchery.pdf}}&2012--2012\tabularnewline
Sacramento R.&SRW&Livingston Stone NFH&Winter&CVWR&111&\raisebox{-0.12 em}{\includegraphics[height=1.02em]{../tex/images/months-SRW_Livingston-Stone-National-Fish-Hatchery.pdf}}&2013--2014\tabularnewline
San Joaquin R.&SJRF&San Joaquin R.&Fall&CVFR&79&\raisebox{-0.12 em}{\includegraphics[height=1.02em]{../tex/images/months-SJRF_San-Joaquin-River.pdf}}&2016--2016\tabularnewline
Trinity R.&TRF&Trinity R. H.&Fall&UKTR&41&\raisebox{-0.12 em}{\includegraphics[height=1.02em]{../tex/images/months-TRF_Trinity-River-Hatchery.pdf}}&2012--2012\tabularnewline
Trinity R.&TRS&Trinity R. H.&Spring&UKTR&46&\raisebox{-0.12 em}{\includegraphics[height=1.02em]{../tex/images/months-TRS_Trinity-River-Hatchery.pdf}}&2012--2012\tabularnewline