### 2024-01-08 lfar-allele-freqs.Rmd

---
title: "Assessing Power of the California Microhap Baseline for GSI (and calculating Fst)"
author: "Eric C. Anderson"
date: "Last Updated: `r Sys.Date()`"
output:
  html_document:
    df_print: paged
    toc: true
    toc_float: true
bibliography: "`r here::here('references.bib')`"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(message = FALSE)
knitr::opts_knit$set(root.dir = here::here())
start_time <- Sys.time()

USING_SNAKEMAKE <- FALSE
NEG_USING_SMK <- TRUE
if(exists("snakemake")) {
  USING_SNAKEMAKE <- TRUE
  NEG_USING_SMK <- FALSE
}
```

```{r rmd_snakemake_hacks, include=FALSE, eval=USING_SNAKEMAKE}
# currently, this block does not get evaluated, while developing,
# but later when we snakemake-ize it all, we will evaluate it.
if(USING_SNAKEMAKE) {
  input_list <- snakemake@input
  output_list <- snakemake@output
}
```

# Input and Output Paths

```{r input_bypass, include=NEG_USING_SMK, eval=NEG_USING_SMK}
# inputs:
input_list <- list(
  final_baseline = "data/cali-chinook-baseline.rds",
  pop_labels = "inputs/reference-collection-names.csv"
)
# outputs:
output_list <- list(
  lfar_tex_table = "tex/inputs/lfar-freqs.tex"
)
# you can create the necessary output directories like this:
dump <- lapply(output_list, function(x)
  dir.create(dirname(x), recursive = TRUE, showWarnings = FALSE)
)
```

# Introduction

The goal here is simply to get the allele frequenccies in different reporting units
at the LFAR markers into a table format.

First, get the baseline and attach the collection and repunit names that we want.
```{r, message=FALSE, warning=FALSE}
library(tidyverse)

# get the full baseline with the new markers, etc.  We will toss out some
# non-winter Sacto fish of unknown provenance
full_base <- read_rds(input_list$final_baseline) %>%
  filter(collection != "SacNonW")

# read the pop labels for proper sorting
pop_labels <- read_csv(input_list$pop_labels)

# rename those collections and repunits
full_base2 <- full_base %>%
  rename(old_name = collection) %>%
  select(-repunit) %>%
  left_join(pop_labels %>% select(-run_timing_group), by = join_by(old_name)) %>%
  select(-old_name) %>%
  select(indiv, repunit, collection, sample_type, everything())

# finally, get the order of collections and repunits we want
repunit_order <- rev(unique(pop_labels$repunit))
```

Now, whittle that down to just the LFAR and pivot it longer:
```{r}
lfar_base <- full_base2 %>%
  select(indiv:collection, starts_with("NC_037130.1")) %>%
  pivot_longer(
    cols = starts_with("NC_037130.1"), 
    names_to = c("locus", "gene_copy"),
    names_pattern = "(NC_037130.1.*)_([12])",
    values_to = "allele"
  )
```

And now we count up the alleles within repunits
```{r}
lfar_cnts <- lfar_base %>%
  filter(!is.na(allele)) %>%
  count(repunit, locus, allele) %>%
  complete(repunit, locus, allele, fill = list(n = 0L)) %>%
  group_by(repunit, locus) %>%
  mutate(
    tot_n = sum(n),
    fract = n / tot_n
  ) %>%
  ungroup()

lfar_cnts
```

Now, we just want to focus on the alleles that are most frequent in the late-fall.
```{r}
focal <- lfar_cnts %>%
  filter(repunit == "CV-Late-Fall") %>%
  group_by(locus) %>%
  filter(n == max(n))

alle_tib <- lfar_cnts %>%
  semi_join(focal, by = join_by(locus, allele)) %>%
  mutate(
    repunit_f = factor(repunit, levels = repunit_order),
    f = sprintf("%.3f", fract)
  ) %>%
  select(-repunit, -allele, -n, -fract) %>%
  pivot_wider(names_from = locus, values_from = c(f, tot_n)) %>%
  arrange(repunit_f) %>%
  select(repunit_f, `f_NC_037130.1:864908-865208`, `tot_n_NC_037130.1:864908-865208`, `f_NC_037130.1:1062935-1063235`, `tot_n_NC_037130.1:1062935-1063235`)

alle_tib
```

Now, we will print this out without any column names so that we can use it in a TeX table.
```{r}
write_delim(alle_tib, delim = "&", eol = "\\tabularnewline\n", file = output_list$lfar_tex_table, col_names = FALSE)
```

# Citations

::: {#refs}
:::

# Session Info

```{r}
sessioninfo::session_info()
```

# Running Time

Running the code and rendering this notebook required approximately this much time:

```{r}
Sys.time() - start_time
```