### 2024-02-14 ckmr-sim-evaluation.Rmd
Working through the CKMRsim stuff.

---
title: "Assessing Power of the California Microhap Baseline for GSI (and calculating Fst)"
author: "Eric C. Anderson"
date: "Last Updated: `r Sys.Date()`"
output:
  html_document:
    df_print: paged
    toc: true
    toc_float: true

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(message = FALSE)
start_time <- Sys.time()
```

```{r}
if(exists("snakemake")) {
  input_list <- snakemake@input
  output_list <- snakemake@output
} else {
  input_list <- list(
    final_baseline = "data/cali-chinook-baseline.rds",
    pop_labels = "inputs/reference-collection-names.csv",
    locus_info = "inputs/ChinookFullPanelLociTable.csv",
    chrom_lengths = "inputs/Otsh_v1.0_chrom_lengths.tsv"
  )
}
```

## Wrangle the data

```{r}
library(tidyverse)
library(CKMRsim)

# get the full baseline with the new markers, etc.  We will toss out some
# non-winter Sacto fish of unknown provenance
full_base <- read_rds(input_list$final_baseline) %>%
  filter(collection != "SacNonW")

# read the pop labels for proper sorting
pop_labels <- read_csv(input_list$pop_labels)

# rename those collections and repunits
full_base2 <- full_base %>%
  rename(old_name = collection) %>%
  select(-repunit) %>%
  left_join(pop_labels %>% select(old_name, collection, repunit, ckmr_sim_group), by = join_by(old_name)) %>%
  select(-old_name) %>%
  select(indiv, repunit, collection, sample_type, ckmr_sim_group, everything())

# after that, we can nest things into separate ckmr_sim_groups
ckmr_sim_nests <- full_base2 %>%
  group_by(ckmr_sim_group) %>%
  nest()

# now, we also want to get the genome positions of things.
# we take its position to be the start position of the amplicon
gen_coords <- read_csv(input_list$locus_info) %>%
  select(Locus, AmpliconGenomeV1Location) %>%
  separate(AmpliconGenomeV1Location, into = c("Chrom", "pos_str"), sep = ":") %>%
  separate(pos_str, into = c("Start", "Stop"), sep = "-", convert = TRUE) %>%
  rename(Pos = Start) %>%
  select(-Stop) %>%
  arrange(Chrom, Pos)

# now we have to do some funny stuff to deal with things on unplaced scaffolds.
# We will just sprinkle scaffolds into the existing chromosome lengths from the
# Otsh_v1.0 genome
chrom_lengths <- read_tsv(input_list$chrom_lengths)

# get the unique unplaced scaffolds and assign them randomly to a position
# in an assembled chromosome.  We set it up so that markers on the same scaffold
# will not extend beyond the end of the chromosome
set.seed(5)
rando_coords <- gen_coords %>%
  filter(str_detect(Chrom, "NW_") | is.na(Chrom)) %>%
  group_by(Chrom) %>%
  mutate(max_pos = max(Pos)) %>%
  select(Chrom, max_pos) %>%
  distinct() %>%
  mutate(
    # this gets the chromosome they are assigned to
    ass_chrom = sample(chrom_lengths$chrom, size = n(), prob = chrom_lengths$num_bases, replace = TRUE)
  ) %>% # now we do some stuff to place the scaffold randomly on the chromosome
  left_join(chrom_lengths, by = join_by(ass_chrom == chrom)) %>%
  mutate(
    ass_pos = floor(runif(n = n(), min = 1, max = num_bases - max_pos - 1 ))
  ) %>%
  select(Chrom, ass_chrom, ass_pos) %>%
  ungroup()

# now we have the information to randomly stick the scaffold markers onto
# chromosomes in the genome.  We do this by joining things onto gen_coords
gen_coords2 <- gen_coords %>%
  left_join(rando_coords, by = join_by(Chrom)) %>%
  mutate(
    Pos = case_when(
      is.na(ass_pos) ~ Pos,
      TRUE ~ ass_pos + Pos
    ),
    ass_chrom = case_when(
      is.na(ass_chrom) ~ Chrom,
      TRUE ~ ass_chrom
    )
  ) %>%
  mutate(Chrom = ass_chrom) %>%
  select(Locus, Chrom, Pos)

```

## Now, a series of functions to operate on each nest and do CKMRsim things

```{r}
#' @param X a tibble like that nested in a single row of ckmr_sim_nests
#' @param GC2 essentially the gen_coords2 object from above
ckmr_afreqs_ready <- function(X, GC2) {
  tmp <- X %>%
    select(-(repunit:sample_type)) %>%
    pivot_longer(
      -indiv,
      names_to = c("Locus", "gene_copy"),
      values_to = "Allele",
      names_pattern = "^(.*)_([12])$"
    ) %>%
    filter(!is.na(Allele)) %>%
    count(Locus, Allele) %>%
    group_by(Locus) %>%
    filter(n() > 1) %>% # filter out monomorphic loci
    mutate(
      Freq = n / sum(n),
    ) %>%
  ungroup() %>%
  left_join(GC2, by = join_by(Locus)) %>%
  select(Chrom, Pos, Locus, Allele, Freq) %>%
  arrange(Pos, desc(Freq))

  no_pos_posse <- tmp %>% 
    filter(is.na(Chrom))

  if(nrow(no_pos_posse) > 0) {
    warning("Some loci without genomic positions: ", paste(unique(no_pos_posse$Locus), collapse = ", "), "\n\n")
    warning("Those loci were dropped, and the analysis should survive, but you should see what's up!")
  }

  # then, return thing without the no_pos_posse markers
  tmp %>%
    filter(!is.na(Chrom)) %>%
    mutate(
      AlleIdx = NA,
      LocIdx = NA
    ) %>%
    CKMRsim::reindex_markers()

}

# here is one to create a ckmr object.
# Y is simply the output of ckmr_afreqs_ready()
ckmr_create <- function(Y) {
  create_ckmr(
    D = Y,
    kappa_matrix = kappas[c("PO", "FS", "HS", "U"), ],
    ge_mod_assumed = ge_model_TGIE,
    ge_mod_true = ge_model_TGIE,
    ge_mod_assumed_pars_list = list(epsilon = 0.005),
    ge_mod_true_pars_list = list(epsilon = 0.005)
  )
}

# This does the simulate_Qij step and then also estimates the FPRs.
ckmr_simQ <- function(C) {
  simulate_Qij(
    C, 
    calc_relats = c("PO", "FS", "HS", "U"),
    sim_relats = c("PO", "FS", "HS", "U"),
    unlinked = FALSE,
    pedigree_list = pedigrees
  )
}

```

This is just here while I develop this:
```{r}
X <- ckmr_sim_nests$data[[5]]
GC2 <- gen_coords2

Y <- ckmr_afreqs_ready(X, GC2)

C <- ckmr_create(Y)

Q <- ckmr_simQ(C)
```