### 2024-03-15 development/genepop-conversion.Rmd
Add genepop conversion to dev directory
Just didnt want to lose it.

---
title: "Genepop Conversion Etc"
output: html_notebook
---

# Making a genepop file

I ended up not using this, but we might want to, ultimately, if we want to look at
LD, etc.

```{r}
gpt <- full_base2 %>%
  select(-repunit, -sample_type) %>%
  pivot_longer(
    cols = -c(indiv, collection),
    names_to = c("locus", "gene_copy"),
    values_to = "allele",
    names_pattern = "(.*)[_.]([12])$"
  ) %>%
  group_by(locus) %>%
  mutate(alle_int = replace_na(as.integer(factor(allele)), 0L)) %>%
  ungroup() %>%
  mutate(alle_str = sprintf("%02d", alle_int)) %>%
  select(-allele, -alle_int) %>%
  pivot_wider(
    names_from = gene_copy,
    values_from = alle_str
  ) %>%
  mutate(geno = str_c(`1`, `2`)) %>%
  select(-`1`, -`2`) %>%
  pivot_wider(
    names_from = locus,
    values_from = geno
  ) %>%
  select(-indiv)

gpt_nest = gpt %>%
  mutate(collection2 = str_c(collection, ","), .after = collection) %>%
  group_by(collection) %>%
  nest()

# now we write that out to a file
outf <- "results/GSI_and_Fst/genepop.txt"
cat("Baseline data set in genepop format\n    ", file = outf)
cat(names(gpt)[-1], sep = ",", file = outf, append = TRUE)
cat("\n", file = outf, append = TRUE)
for(x in gpt_nest$data) {
    cat("POP\n", file = outf, append = TRUE); 
    write_tsv(x, col_names = FALSE, file = outf, append = TRUE)
  }
```