### 2024-06-21 extras/Applying-New-Names-To-Baseline/applying-new-names-to-baseline.Rmd
Beginnging to apply new locus names to the baseline
But I am going to have to get a more complete baseline from Anthony.

---
title: "Applying the new names to the baseline"
output: html_notebook
---

We are just going to take the most recent version of the 
baseline and munge it to change the names of the loci.
```{r}
library(tidyverse)

amplicons <- read_csv("../../inputs/Calif-Chinook-Amplicon-Panel-Information.csv")

new_names <- amplicons %>%
  select(AmpliconName, OtherName)

base <- read_csv("subRoSA_baseline_rubias_JUN2024_w_mixture_complete-rosa_lfar_vgll3six6_added.csv") %>%
  pivot_longer(
    cols = -(indiv:sample_type),
    names_to = c("locus", "gene_copy"),
    values_to = "allele",
    names_pattern = "^(.+)_([12])$"
  ) %>%
  left_join(new_names, by = join_by(locus == OtherName))

```

OK, now, let's check and see which ones don't have new names:

```{r}
base %>%
  filter(is.na(AmpliconName)) %>%
  count(AmpliconName, locus)
```
OK, cool.  Those are the LFAR's that we dropped.  We expect that they won't have new names
because we don't really want them in there.

Let's see if there are any AmpliconName's that don't appear in base.

```{r}
amplicons %>%
  anti_join(base, by = join_by(AmpliconName))
```