### 2024-07-02 extras/Applying-New-Names-To-Baseline/applying-new-names-to-baseline.Rmd
Change sample column name to indiv

---
title: "Applying the new names to the baseline"
output: html_notebook
---

We are just going to take the most recent version of the 
baseline and munge it to change the names of the loci.
```{r}
library(tidyverse)

amplicons <- read_csv("../../inputs/Calif-Chinook-Amplicon-Panel-Information.csv")

new_names <- amplicons %>%
  select(AmpliconName, OtherName)

csv <- read_csv("subRoSA_baseline_rubias_with_rosa_wrap_lfar_vgll3six6_and_mixture_28JUN2024.csv")

# for some reason, there are 22 fish with identical sample_IDs.  
dupies <- csv %>%
  count(Sample_ID) %>%
  filter(n > 1) %>%
  pull(Sample_ID)

# there doesn't seem to be too much any difference in how much missing data each copy of
# each duplicated fish has:
mat <- csv %>% filter(Sample_ID %in% dupies) %>% as.matrix()
rowSums(is.na(mat))

# So, we will just drop the second instance of each:
csv2 <- csv %>%
  filter(!duplicated(Sample_ID))

base <- csv2 %>%
  pivot_longer(
    cols = -(Sample_ID:sample_type),
    names_to = c("locus", "gene_copy"),
    values_to = "allele",
    names_pattern = "^(.+)_([12])$"
  ) %>%
  left_join(new_names, by = join_by(locus == OtherName))

```

Let us count how many things match up here.
```{r}
base %>%
  filter(!is.na(AmpliconName) & !is.na(locus)) %>%
  count(locus, AmpliconName)

```

193 sounds about right.  Those are probably just the RoSAs and the SexID marker.

OK, now, let's check and see which ones don't have new names:

```{r}
base %>%
  filter(is.na(AmpliconName)) %>%
  count(AmpliconName, locus)
```
OK, cool.  Those look like the LFAR's and the WRAPs that we do not include in the
final baseline.  We expect that they won't have new names
because we don't really want them in there.

Let's see if there are any AmpliconName's that don't appear in base.

```{r}
amplicons %>%
  anti_join(base, by = join_by(AmpliconName))
```
Only the RoSA's and the SexID marker.  OK.  That is looking good.

So, let's put it all back together again with the other names.
```{r}
wide <- base %>%
  filter(!is.na(locus) & !is.na(AmpliconName)) %>% # drops the RoSA stuff
  select(-locus) %>%
  pivot_wider(names_from = c(AmpliconName, gene_copy), values_from = allele, names_sep = "_")

# but we want to put things in coordinate sorted order
loci <- amplicons$AmpliconName # get locus names
loci <- loci[!loci == "Ots_sexy001_NW024608692.1_00004816"]  # remove the sexID locus, cuz we don't have it
loci <- loci[!str_detect(loci, "rosa")]  # drop the RoSAs cuz they are in there as the genotype string

loccols <- paste(rep(loci, each = 2), c(1, 2), sep = "_")

wide_ordered_no_rosa <- wide[, c("Sample_ID", "repunit", "collection", "sample_type", loccols)]

# and then we also want to add the rosa genotypes on there, AND remove the SacNonW fish
final <- wide_ordered_no_rosa %>%
  left_join(csv2 %>% select(Sample_ID, hapstr), by = join_by(Sample_ID)) %>%
  mutate(rosa_genotypes_str = hapstr, .after = sample_type) %>%
  select(-hapstr) %>%
  filter(!(collection == "SacNonW")) %>%
  rename(indiv = Sample_ID)
```

Now, we can write that out.

```{r}
dir.create("results", showWarnings = FALSE)
write_csv(final, "results/subRoSA_baseline_rubias_loci_renamed_July_2024.csv.gz")
```