### 2024-07-02 tex/inputs/samples-table.tex

Mainstem River$^1$&Code&Location Name$^2$&Run&ESU$^3$&$N$&$N_\mathrm{day}^4$&Sampling Months&Year Range\tabularnewline
\hline Rogue R.&CRHS&Cole Rivers H.&Spring&SONCC&50&0&\raisebox{-0.12 em}{\includegraphics[height=1.02em]{../tex/images/months-CRHS_Cole-Rivers-Hatchery.pdf}}&2004--2010\tabularnewline
Smith R.&SmRF&Smith R.&Fall&SONCC&46&46&\raisebox{-0.12 em}{\includegraphics[height=1.02em]{../tex/images/months-SmRF_Smith-River.pdf}}&2010--2011\tabularnewline
Klamath R.&BCkF&Blue Ck.&Fall&SONCC&56&56&\raisebox{-0.12 em}{\includegraphics[height=1.02em]{../tex/images/months-BCkF_Blue-Creek.pdf}}&2008--2008\tabularnewline
Klamath R.&IGHF&Iron Gate H.&Fall&UKTR&93&92&\raisebox{-0.12 em}{\includegraphics[height=1.02em]{../tex/images/months-IGHF_Iron-Gate-Hatchery.pdf}}&2017--2017\tabularnewline
Trinity R.&TRS&Trinity R. H.&Spring&UKTR&46&46&\raisebox{-0.12 em}{\includegraphics[height=1.02em]{../tex/images/months-TRS_Trinity-River-Hatchery.pdf}}&2012--2012\tabularnewline
Trinity R.&TRF&Trinity R. H.&Fall&UKTR&41&41&\raisebox{-0.12 em}{\includegraphics[height=1.02em]{../tex/images/months-TRF_Trinity-River-Hatchery.pdf}}&2012--2012\tabularnewline
Eel R.&ERF&Eel R.&Fall&CCC&93&0&\raisebox{-0.12 em}{\includegraphics[height=1.02em]{../tex/images/months-ERF_Van-Arsdale-Fisheries-Station.pdf}}&2014--2014\tabularnewline
Russian R.&RRF&Warm Springs H.&Fall&CCC&92&92&\raisebox{-0.12 em}{\includegraphics[height=1.02em]{../tex/images/months-RRF_Warm-Springs-Hatchery.pdf}}&2012--2012\tabularnewline
Sacramento R.&SRW&Livingston Stone NFH&Winter&CVWR&111&20&\raisebox{-0.12 em}{\includegraphics[height=1.02em]{../tex/images/months-SRW_Livingston-Stone-National-Fish-Hatchery.pdf}}&2013--2014\tabularnewline
Sacramento R.&BCS&Butte Ck.&Spring&CVSR&116&116&\raisebox{-0.12 em}{\includegraphics[height=1.02em]{../tex/images/months-BCS_Butte-Creek.pdf}}&2004--2017\tabularnewline
Sacramento R.&MDS&Deer Ck.&Spring&CVSR&34&34&\raisebox{-0.12 em}{\includegraphics[height=1.02em]{../tex/images/months-MDS_Deer-Creek.pdf}}&2002--2002\tabularnewline
Sacramento R.&MDS&Mill Ck.&Spring&CVSR&50&50&\raisebox{-0.12 em}{\includegraphics[height=1.02em]{../tex/images/months-MDS_Mill-Creek.pdf}}&2002--2004\tabularnewline
Sacramento R.&FRHS&Feather R. H.&Spring&CVSR&202&202&\raisebox{-0.12 em}{\includegraphics[height=1.02em]{../tex/images/months-FRHS_Feather-River-Hatchery.pdf}}&2010--2019\tabularnewline
Sacramento R.&FRHF&Feather R. H.&Fall&CVFR&106&106&\raisebox{-0.12 em}{\includegraphics[height=1.02em]{../tex/images/months-FRHF_Feather-River-Hatchery.pdf}}&2008--2010\tabularnewline
Sacramento R.&BCF&Butte Ck.&Fall&CVFR&45&45&\raisebox{-0.12 em}{\includegraphics[height=1.02em]{../tex/images/months-BCF_Butte-Creek.pdf}}&2004--2016\tabularnewline
Sacramento R.&MDF&Deer Ck.&Fall&CVFR&24&24&\raisebox{-0.12 em}{\includegraphics[height=1.02em]{../tex/images/months-MDF_Deer-Creek.pdf}}&2002--2004\tabularnewline
Sacramento R.&MDF&Mill Ck.&Fall&CVFR&49&49&\raisebox{-0.12 em}{\includegraphics[height=1.02em]{../tex/images/months-MDF_Mill-Creek.pdf}}&2002--2004\tabularnewline
San Joaquin R.&SJRF&San Joaquin R.&Fall&CVFR&79&79&\raisebox{-0.12 em}{\includegraphics[height=1.02em]{../tex/images/months-SJRF_San-Joaquin-River.pdf}}&2016--2016\tabularnewline
Sacramento R.&CHLF&Coleman NFH&Late-fall&CVFR&261&261&\raisebox{-0.12 em}{\includegraphics[height=1.02em]{../tex/images/months-CHLF_Coleman-National-Fish-Hatchery.pdf}}&2011--2015\tabularnewline