### 2024-07-08 extras/Renaming-amplicons/add-species-id.Rmd
Rename markers to include the coho marker in there

---
title: "Renaming all Amplicons to add the Coho-species-id marker"
output: html_notebook
---

After running through the Amplicon-renaming.Rmd notebook, we realized that we
really should add the species ID marker in there.   Doing so wil mean
renumbering all the new names for the scon markers.  No worries.  We can do that.

We will pull it from the renamed input file. Should not be a big deal.

```{r}
library(tidyverse)

amplicons <- read_csv("../../inputs/Calif-Chinook-Amplicon-Panel-Information.csv")

amp2 <- amplicons %>%
  mutate(LocusOrigin = ifelse(OtherName == "OkiOts_120255-113", "Coho", LocusOrigin))

# now we just rename all these guys:
renamed <- amp2 %>%
  mutate(
    tmpname = case_match(
      LocusOrigin,
      "Coho" ~ "coho",
      "LFAR" ~ "lfar",
      "Microhap" ~ "mhap",
      "RoSA" ~ "rosa",
      "SNPtype" ~ "scon",
      "Sex ID" ~ "sexy",
      "Six6" ~ "sixx",
      "VGLL3" ~ "vgll",
      "WRAP" ~ "wrap"
    )
  ) %>% 
  group_by(tmpname) %>%
  mutate(
    AmpliconName = str_c(
      "Ots_", tmpname,
      sprintf("%03d", 1:n()), "_", 
      Otsh_v2.0_chromosome, "_", 
      sprintf("%08d",AmpliconStartPos)),
    .before = Otsh_v2.0_chromosome
  ) %>%
  ungroup() %>%
  select(-tmpname)

# I checked and that is all good with just the change in the coho one (and renumbering).

# So, we can write it out now.
write_csv(renamed, "../../inputs/Calif-Chinook-Amplicon-Panel-Information.csv")
```

Now, let's just go ahead and record the number of amplicons of each type.
```{r}
renamed %>%
  count(LocusOrigin)
```

And make sure that comports with how we have named them all:
```{r}
renamed %>%
  extract(AmpliconName, into = "categ", regex = "Ots_([a-z]+).*$") %>%
  count(categ)
```

Yep.  That is all good.