### 2024-07-18 006-winter-run-vs-all-cv-alle-freqs.Rmd
Getting WRAP stuff added in to repo and paper/supp

---
title: "Absolute Value of Alle Freq Diff between Winter Run and all other Central Valley fish"
biblio-style: apalike
output: 
  html_document:
    df_print: paged
    toc: true
    toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(message = FALSE)
knitr::opts_knit$set(root.dir = here::here())
start_time <- Sys.time()
```

# Input and Output Paths

```{r}
if(exists("snakemake")) {
  input_list <- snakemake@input
  output_list <- snakemake@output
} else {
  # inputs:
  input_list <- list(
    wrap_chr_len = "inputs/wrap/chrom_lengths.txt",
    win_non_win_abs_diffs = "stored_results/win-non-win-abs-diffs-gt0.5.rds"
  )
  # outputs:
  output_list <- list(
    mh_plot = "results/wrap-discover/winter-v-non-winter-mh-plot_nind_ge10.pdf",
    mh4tex = "tex/images/winter-v-non-winter-mh-plot_nind_ge10.pdf",
    pdf = "results/genomic-locations-plot/gen-loc.pdf",
    hundy_kb_plot = "results/wrap-discover/100-kb-intervals-plot.pdf",
    hundy_kb4tex = "tex/images/wrap-slide-window.pdf",
    fuv2 = "results/wrap-discover/follow_up_variants2.rds",
    wrap61_plot = "results/wrap-discover/wrap-candi.pdf",
    wrap61_4tex = "tex/images/wrap-candi.pdf"
  )
}
# you can create the necessary output directories like this:
dump <- lapply(output_list, function(x)
  dir.create(dirname(x), recursive = TRUE, showWarnings = FALSE)
)
```

# Stuff Run on the Cluster

## Computing the allele frequencies across the whole genome

### Packages and paths 
```r
library(tidyverse)
library(viridis)
dir.create("outputs/501", recursive = TRUE, showWarnings = FALSE)
dir.create("intermediates/501", recursive = TRUE, showWarnings = FALSE)
```

## Text files with names of fall-run and spring-run fish

We do this in R.
```r
pmeta <- read_csv("data/wgs-chinook-samples.csv")

# get file with fall-run names in it
pmeta %>%
  filter(run_type == "Winter") %>%
  .$vcf_name %>%
  cat(., file = "intermediates/501/winter-run.txt", sep = "\n")

# get file with spring-run names in it
pmeta %>%
  filter(
    !str_detect(Population, "^Salmon|^Trinity"),
    run_type != "Winter") %>%
  .$vcf_name %>%
  cat(., file = "intermediates/501/non-winter-run-cv.txt", sep = "\n")
```

## Clean up the damn VCF files

Why is this still a problem?  Three of the VCF files have non-ascii gremlins in them.
We clean them up with `tr` as shown below. Doing so gave us the correct md5 hashes for the
files when we got them from Russ.  But apparently, the fixed versions were never put back
up onto Google Drive.  WTF?!  I will fix and do that now...
```sh
conda activate bioinf # to get bgzip
cd chinook_WGS_processed
for i in NC_037097.1.vcf.gz NC_037109.1.vcf.gz NC_037116.1.vcf.gz; do 
  echo $i;
  zcat $i | tr -cd '\11\12\15\40-\176' | bgzip > cleaned_$i; 
done

# when done:
for i in cleaned_NC_037*.vcf.gz; do j=${i/cleaned_/}; echo $i $j; mv $i $j; done

# after I had indexed all the VCFs I then rcloned everything back
# to the Google Drive
rclone copy --drive-shared-with-me chinook_WGS_processed gdrive-rclone:chinook_WGS_processed

```

## Do it all in a job array

**NOTE!!** I installed angsd 0.921 with conda because the latest version
fails when parsing the VCFs.

First, some prep:
```sh
# get a list of the vcf files to process
ls -l chinook_WGS_processed/*.vcf.gz | awk 'BEGIN{printf("index\tvcf\n");} {printf("%d\t%s\n", ++n, $NF);}' > intermediates/501/vcflist.txt

# get some directories where we need them:
mkdir -p intermediates/501/slurm_{out,err}
```

Then we make a quick job array script called `script/501-winter-non-winter-alle-freqs.sh`
and run it.

First test it on the first 3:
```sh
sbatch --array=1-3 script/501-winter-non-winter-alle-freqs.sh
```
That seems successful, so we crank it out on the remining ones
```sh
sbatch --array=4-41 script/501-winter-non-winter-alle-freqs.sh
```

## Filtering, alle freq diffs, and plotting

I am still doing this on the cluster within R. Our strategy
is to read the winter run and the non-winter-run-cv separately, and filter each
on the read depth requirement, separately.  Then we will do an inner join of all those.

I am requiring 62.5% or more having at least one read. So that is 10/16 for the
winter run and 50/80 for the nonwinter run.

Get the winter run:
```r
winter_files <- dir(
  path = "intermediates/501/winter-run",
  pattern = "*mafs.gz",
  full.names = TRUE
)

winter_freqs <- lapply(winter_files, function(x) {
  read_tsv(x) %>%
    filter(nInd >= 10)
}) %>%
  bind_rows()
```

Then get the non-winter-run:
```r
nonwinter_files <- dir(
  path = "intermediates/501/non-winter-run-cv",
  pattern = "*mafs.gz",
  full.names = TRUE
)

nonwinter_freqs <- lapply(nonwinter_files, function(x) {
  read_tsv(x) %>%
    filter(nInd >= 50)
}) %>%
  bind_rows()
```
Now we do the inner join:
```r
all_them <- inner_join(
  winter_freqs,
  nonwinter_freqs,
  by = c("chromo", "position"),
  suffix = c("_winter", "_nonwinter")
) %>%
  mutate(
    ave_freq = (unknownEM_winter + unknownEM_nonwinter) / 2,
    abs_diff = abs(unknownEM_winter - unknownEM_nonwinter)
  )
```
And it is quite straightforward from here. Note that we have 7,295,001 SNPs here.

First, just look at the distribution of absolute differences:
```r
g <- ggplot(all_them, aes(x = abs_diff)) +
  geom_histogram(binwidth = 0.001)
g
```
That looks good.  What if we focus on everything with an abs diff > 0.25?
```r
g +
  xlim(0.25, 1.01)
```
There is a gradual decline toward 1.00, but a tiny bump up at 1.00 itself.

This is going to be hard...
```r
g +
  xlim(0.8, 1.01)
```

Have a look at everything > 0.95 with rView:
```r
library(rView)
all_them %>%
  filter(abs_diff > 0.95) %>%
  arrange(desc(abs_diff)) %>%
  select(chromo, position, abs_diff, everything()) %>%
  count(chromo) %>%
  arrange(desc(n)) %>%
  rView()
```

So, I think what I want to do now is put everything > 0.5 into stored_results.
That still gives us 175 K SNPs to plot.
```r
dir.create("stored_results/501", showWarnings = FALSE, recursive = TRUE)
all_them %>%
  filter(abs_diff >=  0.5) %>%
  write_rds("stored_results/501/win-non-win-abs-diffs-gt0.5.rds", compress = "xz")
```

# Back to the laptop

## "Manhattan" plot

I have some code for this, but need the chromosome lengths:
We need to get the chromosome lengths for all of these.  We can pull those out of the
VCF file.
```sh
mkdir -p intermediates/501
mkdir -p outputs/501
(echo "Chromosome chrom_length"; gunzip -c data/greb1l-ish-region.vcf.gz 2> /dev/null | awk -F"=" '/^##contig/ {print $3, $4} !/^#/ {exit}' | sed 's/,length//g; s/>//g;') > inputs/wrap/chrom_lengths.txt
```

Now, filter down to abs_diff > 0.25, and get the chrom_lengths
```{r}
library(tidyverse)

wnw_lite <- read_rds(input_list$win_non_win_abs_diffs) %>%
  rename(Chromosome = chromo)

chrom_lengths <- read_table(input_list$wrap_chr_len)
```

Now, my function to make the plot. This website [https://www.r-graph-gallery.com/wp-content/uploads/2018/02/Manhattan_plot_in_R.html](https://www.r-graph-gallery.com/wp-content/uploads/2018/02/Manhattan_plot_in_R.html) had a nice discussion of it. We've store
two functions in `R/`: `my_mh_prep` and `plot_mh` and source them here: 
```{r}
source("R/manhattan_plot_funcs.R")
```

Now, use those funcs:
```{r}
wnw_prepped <- my_mh_prep(wnw_lite, chrom_lengths)

mh_plot <- plot_mh(wnw_prepped$snps, wnw_prepped$axisdf) +
  xlab("Position along chromosome") +
  ylab("Absolute value of allele frequency difference, spring-run vs. fall-run")

mh_plot
```

So, as expected, there are a lot of things with large allele frequency differences.  And there are
those gaps, from before, too.  (Places where the VCFs got screwed up, somehow).

Let us also be clear that some of these SNPs might have only 10 individuals, each with just one read.  So, even though they are diploids, that is still just a sample of 10 gene copies---not 20.  Let's try filtering to 
16 individuals.

First, just look at the distribution of number of winter indivs with reads:
```{r}
wnw_lite %>%
  count(nInd_winter)
```

Plot it with sites with at least 14.
```{r}
wnw_prepped_filt <- my_mh_prep(wnw_lite %>% filter(nInd_winter >= 14), chrom_lengths)

mh_plot_filt <- plot_mh(wnw_prepped_filt$snps, wnw_prepped_filt$axisdf) +
  xlab("Position along chromosome") +
  ylab("Absolute value of allele frequency difference, spring-run vs. fall-run")

mh_plot_filt
```

So, the standouts there are still interesting.

Hell, do it with 16, too:
```{r}
wnw_prepped_filt <- my_mh_prep(wnw_lite %>% filter(nInd_winter >= 16), chrom_lengths)

mh_plot_filt <- plot_mh(wnw_prepped_filt$snps, wnw_prepped_filt$axisdf) +
  xlab("Position along chromosome") +
  ylab("Absolute value of allele frequency difference, spring-run vs. fall-run")

mh_plot_filt
```

Which confirms that the NC_037112.1 chromosome is the most interesting.

Let's plot a big long version of the full plot so we can 
see the points and point density better.
```{r}
ggsave(mh_plot + ylim(0.5, 1.0), filename = output_list$mh_plot,
       height = , width = 12)

file.copy(from = output_list$mh_plot, to = output_list$mh4tex, overwrite = TRUE)
```

## Zoom in on NC077112.1

```{r}
wnw_lite %>%
  filter(Chromosome == "NC_037112.1", position > 2.37e7, position < 2.65e7) %>%
  ggplot(aes(x = position, y = abs_diff)) + 
  geom_point()
```

Let's also just do a smooth for average number of sites with abs_diff > .9 per megabase.
```{r}
hundy_kb_chunks <- wnw_lite %>%
  select(Chromosome, position, abs_diff) %>%
  mutate(
    geq90 = abs_diff >= 0.9
  ) %>%
  group_by(Chromosome) %>%
  mutate(length = max(position) - min(position)) %>%
  dplyr::do(
    as_tibble(
      ksmooth(
        .$position,
        .$geq90,
        bandwidth = 1e6,
        n.points = 10 * (1 + .$length[1] / 1e6)
      )
    )
  ) %>%
  ungroup() %>%
  rename(
    `100kb_midpoint` = x,
    fract_snps_with_abs_diff_geq_0.9 = y  
  )

hundy_kb_chunks %>%
  arrange(desc(fract_snps_with_abs_diff_geq_0.9 )) %>%
  slice(1:20)

```

That is pretty compelling.  There a 1 Mb chunk on 37112 that is more interesting than anythign else.
Let's make a smoothed Manhattan plot of this:
```{r}
hundy_prepped <- hundy_kb_chunks %>%
  mutate(
    position = floor(`100kb_midpoint`),
    abs_diff = fract_snps_with_abs_diff_geq_0.9
  ) %>% 
  my_mh_prep(chrom_lengths)

hundy_mh_plot <- plot_mh(hundy_prepped$snps, hundy_prepped$axisdf) +
  xlab("Midpoint of 100 Kb interval") +
  ylab("Fraction of sites with abs_diff > 0.5 that have abs_diff >= 0.9") +
  ylim(0,NA) +
  geom_line(aes(color = color_band, group = Chromosome), size = 0.4) +
  geom_point(aes(color = color_band), size = 0.9)

hundy_mh_plot
```

## Investigate SNPs in the "six peaks"

We are going to look at SNPs in the 6 highest peaks from our smoothed 100 Kb windows,
with the possibility of designing some primers for them.

We will define those peaks as places where the smoothed fraction of sites > 0.9 is
greater than 0.124
```{r}
# filter it down and break them into groups of the different "adjacent" chunks
follow_up_chunks <- hundy_kb_chunks %>%
  filter(fract_snps_with_abs_diff_geq_0.9 > 0.124) %>%
  mutate(
    dist = `100kb_midpoint` - lag(`100kb_midpoint`, default = 200000),
    group = cumsum(Chromosome != lag(Chromosome, default = "bonkers") | dist > 100000), # this breaks the second peak on Chr 16 into two, so we will merge them
    group = case_when(
      group == 6 ~ 5L,
      group > 6 ~ group - 1L,
      TRUE ~ group
    )
  )

follow_up_chunks
```

That gives us 6 groups of adjacent (more or less) 100 Kb windows. Let us plot
those out:
```{r}
fups <- hundy_prepped$snps %>% semi_join(follow_up_chunks %>% filter(Chromosome != "NC_037103.1"), by = join_by(Chromosome, `100kb_midpoint`))

red_balloons <- hundy_mh_plot +
  geom_point(data = fups, shape = 21, fill = NA, colour = "red", size = 2)

red_balloons
```

While we are at it, let's put a copy of that into outputs
```{r}
ggsave(red_balloons, filename = output_list$hundy_kb_plot, width = 10, height = 6)

file.copy(from = output_list$hundy_kb_plot, to = output_list$hundy_kb4tex, overwrite = TRUE)
```

I want to plot the
estimated allele freq diffs from those windows (plus a buffer of 70 Kb on either side)
in a big facet_wrap.  
```{r}
follow_up_variants <- follow_up_chunks %>%
  group_by(group) %>%
  summarise(
    Chromosome = Chromosome[1],
    min_pos = min(`100kb_midpoint`) - 5e4,
    max_pos = max(`100kb_midpoint`) + 5e4
    ) %>%
  left_join(wnw_lite, by = "Chromosome") %>%
  filter(position >= min_pos - 7e4, position <= max_pos + 7e4) 
```

With that, we can then plot things. I will color by the number of winter run that
we obtained reads from:
```{r, fig.height=10, fig.width=8}

group_facet_inds <- ggplot(follow_up_variants, aes(x = position, y = abs_diff)) +
  facet_wrap(~ group, scales = "free_x") +
  geom_point(aes(colour = nInd_winter)) +
  scale_color_viridis_c()

# make another showing the SNPs to design on.  That will be: anything above 0.975
# absolute difference, or the highest abs diff, up to 8 variants.
fuv2 <- follow_up_variants %>%
  ungroup() %>%
  arrange(group, desc(abs_diff)) %>%
  group_by(group) %>%
  mutate(
    abs_diff_rank = 1:n(),
    try_design = abs_diff_rank <= 8 | abs_diff > 0.975
  )

group_facet_selecto <- ggplot(fuv2, aes(x = position / 1e06, y = abs_diff)) +
  facet_wrap(~ Chromosome, scales = "free_x") +
  geom_point(aes(fill = try_design), shape = 21, stroke = 0.05) +
  scale_fill_manual(values = c("blue", "red")) +
  xlab("Position on Chromosome (in megabases)") +
  ylab("Absolute difference in allele frequency between\nwinter-run and non-winter-run")

group_facet_selecto
```

How many is that?
```{r}
sum(fuv2$try_design)
```

Let's make a plot of that for the supplement.
```{r}
ggsave(group_facet_selecto, filename = output_list$wrap61_plot, width = 9, height = 6)

file.copy(from = output_list$wrap61_plot, to = output_list$wrap61_4tex, overwrite = TRUE)
```

## Finally, save fuv2 for later use

```{r}
write_rds(fuv2, path = output_list$fuv2, compress = "xz")
```

# Citations

::: {#refs}
:::

# Session Info

```{r}
sessioninfo::session_info()
```

# Running Time

Running the code and rendering this notebook required approximately this much time:

```{r}
Sys.time() - start_time
```