### 2024-08-19 tex/inputs/ascertainment-samples-supp-table.tex
Add ascertainment panel sample size table to supp

Sacramento River winter run&4 \\
Feather River spring run&2 \\
Feather River fall run&2 \\
Butte Creek spring run&2 \\
Deer Creek fall run&2\\
Eel River fall run&2 \\ 
Klamath Basin, Trinity River Hatchery spring run&2\\
Southern Oregon, Chetco River&2\\
Northern Oregon, Siletz River&2\\
Columbia River, Kalama Hatchery spring run&2\\
Upper Columbia River, Wenatchee River&2\\
Snake River, McCall Hatchery summer-spring run&2\\
Puget Sound,Kendall Hatchery spring run&2\\
Fraser River, Thompson River, Spius Creek&2\\
Fraser River, Birkenhead River&2\\