### 2024-08-26 007-whoa-on-chinook.Rmd
Commit the Whoa stuff

---
title: "Running the California Microhap Baseline through `whoa` and Deciding We Should Drop Two of the Amplicons"
author: "Eric C. Anderson"
date: "Last Updated: `r Sys.Date()`"
output:
  html_document:
    df_print: paged
    toc: true
    toc_float: true

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(message = FALSE)
knitr::opts_knit$set(root.dir = here::here())
start_time <- Sys.time()

```

# Input and Output Paths

```{r}
if(exists("snakemake")) {
  input_list <- snakemake@input
  output_list <- snakemake@output
} else {
  input_list <- list(
    final_baseline = "data/SWFSC-chinook-reference-baseline.csv.gz",
    pop_labels = "inputs/reference-collection-names.csv"
  )
  # outputs:
  output_list <- list(
    zs = "results/whoa/heterozygote-z-scores.pdf"
  )
}

# you can create the necessary output directories like this:
dump <- lapply(output_list, function(x)
  dir.create(dirname(x), recursive = TRUE, showWarnings = FALSE)
)
```

```{r, message=FALSE, warning=FALSE}
library(tidyverse)
library(whoa)

# read in the full baseline
full_base <- read_csv(input_list$final_baseline)

# capture that rosa hapstr with some meta data from future use
rosa_genos0 <- full_base %>%
  select(indiv:rosa_genotypes_str)

# read the pop labels for proper sorting
pop_labels <- read_csv(input_list$pop_labels)

# rename those collections and repunits
full_base2 <- full_base %>%
  rename(old_name = collection) %>%
  select(-repunit) %>%
  left_join(pop_labels %>% select(old_name, collection, repunit), by = join_by(old_name)) %>%
  select(-old_name) %>%
  select(indiv, repunit, collection, sample_type, everything()) %>%
  filter(!duplicated(indiv))  # necessary cuz clemento stuck mill and deer together into a single code

# finally, get the order of collections and repunits we want
collection_order <- unique(pop_labels$collection)

```

Get things into long format:
```{r}
fb_long <- full_base2 %>%
  select(-sample_type, -rosa_genotypes_str) %>%
  pivot_longer(
    cols = -(indiv:collection), 
    names_to = c("locus", "gene_copy"),
    values_to = "allele",
    names_pattern = "^(.*)_([12])$"
  )
```

Now, for every allele in the data set we define a new locus---one in which the
allele is the 1 allele and everything else is the 0 allele.

For the purposes of passing things off to whoa, we
are going to use as "loki" the combination of `locus` and `recoding`, so we will
just paste those with a dot.

Whoa says it wants loci named by Chrom-Pos but I don't think that is actually
necessary, so I will just paste recoding on with a dot to the locus name.

```{r}
RecodedAlleles <- fb_long %>%
  filter(!is.na(allele)) %>%   # it is very important to remove missing genotypes
  filter(locus != "Ots_coho001_05_32691399") %>%  # no reason to have the coho allele
  group_by(locus) %>%
  mutate(
    alle_int = as.integer(factor(allele, levels = sort(unique(allele)))),
    alle_cnt = length(levels(factor(allele)))
  ) %>%
  ungroup() %>%
  mutate(
    boing = map2(.x = alle_int, .y = alle_cnt, .f = function(x, y) {
      list(
        recoding = 1:y,
        new_alle = ifelse(x == 1:y, 1L, 0L)
      )
    })
  ) %>%
  mutate(
    recoding = map(boing, "recoding"),
    new_alle = map(boing, "new_alle")
  ) %>%
  select(-boing) %>%
  unnest(cols = c(recoding, new_alle)) %>%
  #extract(locus, into = c("chrom", "pos"), regex = "Ots_[a-z]+[0-9]+_([0-9]+)_([0-9]+)", remove = FALSE) %>%
  #mutate(loki = str_c("Ots", chrom, "_", recoding, "-", pos ), .after = locus) %>%
  mutate(loki = str_c(locus, recoding, sep = "."))
  #select(-chrom, -pos)
```

That seemed to go well.  In order to be able to get back from the new_alle and
recoding to the actual allele, in case we want to, and to verify that things worked,
we can collapse this into the different cases:
```{r}
haplo_integers <- RecodedAlleles %>%
  distinct(locus, loki, allele, alle_int, alle_cnt, recoding, new_alle) %>%
  arrange(locus, recoding, allele)

haplo_integers
```

That all looks just right.

Recall, for whoa we want an 012 matrix with missing data denoted by -1.

So, we need to turn these thing back into genotypes for the different
individuals.  Once I am done with that, I am going to nest the 012 matrices
up by repunit and collection

```{r}
Recoded_012_full_tib <- RecodedAlleles %>%
  select(repunit, collection, indiv, loki, new_alle) %>%
  group_by(repunit, collection, indiv, loki) %>%
  summarise(geno = sum(new_alle)) %>%
  ungroup() %>%
  pivot_wider(names_from = loki, values_from = geno, values_fill = -1L) %>%
  nest(tibble = -c(repunit, collection)) %>%
  mutate(
    d012 = map(tibble, function(x) x %>% select(-indiv) %>% as.matrix() %>% t())
  )
```

Cool, now we have a list column with a d012 for each collection.

Now we can make a new list column with the whoa output.
```{r}
tib_with_plots <- Recoded_012_full_tib %>%
  mutate(
    exp_and_obs = map(d012, function(x) whoa::exp_and_obs_geno_freqs(d012 = x)),
    scatter = pmap(
      list(eo = exp_and_obs, col = collection, ti = tibble),
      function(eo, col, ti) {
        whoa::geno_freqs_scatter(eo, max_plot_loci = 1000) +
          ggtitle(str_c(col, "  (", nrow(ti), ")"))
      }
    ) 
  )
```

We can see those all in the notebook like this
```{r}
tib_with_plots$scatter
```

There are clearly a few alleles that do not show up in heterozygous
form as freuquently as expected.  I suspect that these may belong
to just one or two loci.

To find which loci are involved, we will simply sort on the z-scores
for the heterozygotes.
```{r}
sorted_losers <- tib_with_plots %>%
  select(repunit, collection, exp_and_obs) %>%
  unnest(exp_and_obs) %>%
  filter(geno == "1") %>%
  arrange(z_score)

```

Looking at that, it is clear that two amplicons are heavily implicated: `Ots_mhap073_19_03686095`
and `Ots_mhap093_27_16923862`.  In the Eel river we also see `Ots_mhap053_11_42403691` pop up
once or twice amongst the most extremely low z-scores for hets, and also
we see `Ots_mhap057_13_38673256.2` at Cole Rivers.  Eventually with z-scores above
-3 we start to see more loci that are suspect are just due to random variation.

But, let's investigate those few loci that we saw in the context of all of the loci.
This makes a massive faceted thing that is really only viewable as a PDF, so we just
save it.
```{r}
zs <- sorted_losers %>%
  extract(snp, into = "locus", regex = "^(.+)\\.[0-9]+", remove = FALSE) %>%
  ggplot(aes(x = z_score)) +
  geom_vline(xintercept = 0, color = "red") +
  geom_histogram() +
  theme_bw() +
  facet_wrap(~ locus, ncol = 10)

  ggsave(zs, filename = output_list$zs, width = 20, height = 30)
```

That provides a lovely perspective on things.  mhap073 and mhap093 are clearly pathological.
mhap053 and mhap057 show a slight downward trend in the z-scores, but it must be somewhat
variable by population, and I don't think it is enough to toss them out.

mhap073 and mhap093 have 4 and 2 alleles respectively, so tossing them won't be
a big deal.  mhap093 only has two alleles but the are named things like
`CCGCCGTTGCT` and `CTGCCGTTGCT`, so that is pretty messed up anyway.

So, I think we should toss those two loci out.

If we do toss them out, this is what the geno freq plots look like.
```{r}
tib_with_dropped_plots <- tib_with_plots %>%
  mutate(dropped_eo = map(exp_and_obs, function(x) x %>% filter(!str_detect(snp, "Ots_mhap073_19_03686095|Ots_mhap093_27_16923862")))) %>%
  mutate(
    dropped_scatters = pmap(
      list(eo = dropped_eo, col = collection, ti = tibble),
      function(eo, col, ti) {
        whoa::geno_freqs_scatter(eo, max_plot_loci = 1000) +
          ggtitle(str_c(col, "  (", nrow(ti), ")"))
      }
    ) 
  )

tib_with_dropped_plots$dropped_scatters
```

So, what we see is that most of the really egregious ones are gone, but in some
populations there are a few that look a little wonky, but those are only
in a few pops.  So, I don't think we really need to toss any others.