### 2025-01-07 tex/data-accessibility.tex
Tweaks.  Mostly anonymizing for review

\section*{Data Accessibility}

All data and code needed to reproduce the results here are available online.
\begin{itemize}
\item Online version of data and scripts used in paper, anonymized for review: \\
          \url{https://anonymous.4open.science/r/california-chinook-microhaps-A661}
        %\url{https://github.com/eriqande/california-chinook-microhaps}
\item Archived version of data and scripts used in paper: \url{zenodo link when made}.
\item Data repository with full baseline data set. UPDATED WHEN JOURNAL PROVIDES DOI. \url{https://dryad.something.or.other}
\end{itemize}
\mbox{}