### 2025-03-14 008-lfar-gwas-and-markers.Rmd
Add lfar markers notebook

---
title: "Late Fall Associated Region: Genome Wide Association Study and Marker Selection"
author: "Eric C. Anderson"
date: "Last Updated: `r Sys.Date()`"
output:
  html_document:
    df_print: paged
    toc: true
    toc_float: true
---

# Introduction

Here we do a GWAS to find a genomic region that is substantially differentiated between
the late-fall and fall runs in the Central Valley, and then we identify several
candidate sites to target for amplicon sequencing.

# Input and Output Paths

```{r}
start_time <- Sys.time()

# inputs:
input_list <- list(
  chr34_late_fall_mafs = "stored_results/5Mb_Chr34-late-fall.mafs.gz",
  chr34_cv_fall_mafs = "stored_results/5Mb_Chr34-cv-fall.mafs.gz",
  gwas_lf_v_sanjo = "stored_results/lfar-gwas/lf_v_sanjo-NC_037130.1.lrt0.gz",
  gwas_lf_v_frhf = "stored_results/lfar-gwas/lf_v_frhf-NC_037130.1.lrt0.gz",
  ots34_snpeff = "stored_results/thompson2020-vcf/ots34-5Mb-snpEff.vcf.gz",
  pvals = "stored_results/stored_p_vals.rds"
)
# outputs:
output_list <- list(
  late_fall_ids = "outputs/007/late-fall-ids.txt",
  cv_fall_ids = "outputs/007/cv-fall-ids.txt",
  candidates_for_assays = "outputs/007/candidates_for_assays.csv"
)
# you can create the necessary output directories like this:
dump <- lapply(output_list, function(x)
  dir.create(dirname(x), recursive = TRUE, showWarnings = FALSE)
)
```

# GWAS

This is all executed with a Snakefile and associated files that reside in the
repository at: `SnakemakeSubDirs/cga-late-fall-bam-assoc`.  The contents of that
directory were originally part of another GitHub repo---simply `eriqande/cga-late-fall-bam-assoc`---but
have been inserted here for ease of archiving.

The figure created by that Snakefile is included in the paper.  We also have saved
the likelihood ratios for the various markers on Chromosome 34 (NC_037130.1)
for both the Late-fall vs San Joaquin fall
and the Late-Fall vs Feather River fall comparisons in `stored_results` so that we can have
those when visualizing allele frequencies.

Additionally, information for all markers with a p-value > 3 in both the SanJo-vs-LateFall and 
the FeatherRiverHatcheryFall-vs-LateFall comparisons was saved in `stored_results/stored_p_vals.rds`
for use now.

## Investigate GWAS pvalues

```{r}
library(tidyverse)
library(vcfR)

# read in the pvals and label the comparisons FRFH and SanJo and then pivot that 
# so we can compare values from the two analyses at each marker
pvals <- read_rds(input_list$pvals) %>%
  mutate(comparison = ifelse(comparison == "Late Fall versus Feather River Hatchery Fall", "FRHF", "SanJo")) %>%
  select(-(Major:Frequency), -(mh_label:xpos), -cgroup) %>%
  pivot_wider(
    names_from = comparison,
    values_from = c(LRT, neg_log10_p)
  )

# here are the total number of SNPs in each comparison:
FRH_snps <- 13477607
SJ_snps <- 12752546

# now, we can look for those that have have log10 pvals in both of the comparisons.
# maybe just plot them first:
ggplot(pvals, aes(x = neg_log10_p_FRHF, y = neg_log10_p_SanJo)) +
  geom_point() +
  geom_hline(yintercept = -log10(5e-8), linetype = "dashed") +
  geom_vline(xintercept = -log10(5e-8), linetype = "dashed")
```

So, that is only a single SNP in both:
```{r}
pvals %>%
  filter(neg_log10_p_FRHF > -log10(5e-8), neg_log10_p_SanJo > -log10(5e-8))
```

Now, let's consider a Bonferroni correction:
```{r}
pv2 <- pvals %>%
  mutate(
    bonFRHF = (10 ^ -neg_log10_p_FRHF) * FRH_snps,
    bonSJ = (10 ^ -neg_log10_p_SanJo) * SJ_snps,
  ) %>%
  arrange(bonFRHF)

pv2
```

So, that is only a single SNP that would be "significant" under a Bonferroni correction,
in the FRFH comparison.

Likewise only a single SNP is "significant" at a Bonferroni corrected 0.05 level
in the SJ comparison, hich we show by sorting the above table on the SanJo
values:
```{r}
pv2 %>%
  arrange(bonSJ)
```

# Marker Selection

This was originally done in
`/Users/eriq/Documents/git-repos/further-chinook-wgs/101-design-late-fall-assays.html`
and it has been transferred here, now.

Mac originally ran the association analysis between Late Fall and the two fall run groups in
the Central Valley (Feather and San Joaquin), finding significant differences
on Chromosome 34 (NC_037130.1).   We have seen the same thing with the GWAS just described.

Accordingly, we will investigate a 5 Mb chunck of Chromosome 34 to:

1. confirm those markers and explore patterns of differentiation in the region
2. See if the differentiation corresponds to annotated variants of large effect
3. identify candidate markers that we should be able to type in many more individuals.
We will provide sequences flanking these candidates so that
Anthony can design primers for them.

## Overview of allele frequency differences in the region

We will simply compare the allele frequencies estimated in the CV fall-run vs the late-fall
run using ANGSD.  I will do this in a 5 Mb chunk for a little context.

### Get the IDs.
```{r, message=FALSE}

# grab the meta data from the Science paper off of GitHub 
nooks <- read_csv("https://raw.githubusercontent.com/eriqande/thompson-et-al-2020-chinook-salmon-migration-timing/master/data/wgs-chinook-samples.csv")

# write out the names of the late-fall fish
nooks %>%
  filter(Population == "Coleman Hatchery Late Fall") %>%
  pull(vcf_name) %>%
  cat(., sep = "\n", file = output_list$late_fall_ids)

# same for the fall-run fish in the CV:
nooks %>%
  filter(Population %in% c("Feather River Hatchery Fall", "San Joaquin River Fall")) %>%
  pull(vcf_name) %>%
  cat(., sep = "\n", file = output_list$cv_fall_ids)
```

### Extract 5 Mb and the two sets of indivs

We show the code but don't actually do it, here.

```{sh, eval=FALSE}
bcftools view -S outputs/007/late-fall-ids.txt -Oz stored_results/thompson2020-vcf/NC_037130.1.vcf.gz NC_037130.1:1-5000000 > outputs/007/late-fall.vcf.gz
bcftools view -S outputs/007/cv-fall-ids.txt -Oz stored_results/thompson2020-vcf/NC_037130.1.vcf.gz NC_037130.1:1-5000000 > outputs/007/cv-fall.vcf.gz
```

### Run ANGSD to estimate allele frequencies

Here is the code used.  We don't do it here, because in and update from angsd 0.920
to 0.938 it broke.  But I have the estimated allele freqs in stored results.

```{sh, eval=FALSE}
cd outputs/007
angsd -vcf-gl late-fall.vcf.gz  -fai ../../inputs/Otsh_v1.0_genomic.fna.fai -nind 16 -domaf 3 -out late-fall 2> late-fall.redirect.stderr
angsd -vcf-gl cv-fall.vcf.gz  -fai ../../inputs/Otsh_v1.0_genomic.fna.fai -nind 32 -domaf 3 -out cv-fall 2> cv-fall.redirect.stderr
```

### Read the estimated allele freqs in and analyse them

```{r, message=FALSE}
af1 <- list(
  late_fall = read_tsv(input_list$chr34_late_fall_mafs),
  cv_fall = read_tsv(input_list$chr34_cv_fall_mafs)
) %>%
  bind_rows(.id = "run") %>%
  pivot_wider(names_from = run, values_from = c(knownEM, unknownEM, nInd)) %>%
  mutate(abs_diff = abs(knownEM_late_fall - knownEM_cv_fall))
```

Let's only bother with ones that have 8 or more indivs with reads in late fall and 16 or more in cv-fall, and
then we can plot them.
```{r}
af2 <- af1 %>%
  filter(nInd_late_fall >= 8 & nInd_cv_fall >= 16)

ggplot(af2, aes(x = position, y = abs_diff)) +
  geom_point(alpha = 0.4)
```

That is pretty cool.  It shows that characteristic pattern near 800 Kb that we saw around GREB1L,
where there a lots of SNPs with large allele frequency differences, and lots with very small
allele frequency differences, but very few with intermediate differences.

One thing I want to confirm is that all the SNPs with small allele freq differences
in that region are actually at low frequency in both groups.  We can do that by
coloring the points by the max frequency (between the two groups) of the minor
allele---that is the allele that is at lowest _average_ frequency in the two populations.

```{r}
af3 <- af2 %>%
  mutate(is_maf = (knownEM_cv_fall + knownEM_late_fall) < 0.5,
         max_maf = pmax(ifelse(is_maf, knownEM_late_fall, 1 - knownEM_late_fall),
                        ifelse(is_maf, knownEM_cv_fall, 1 - knownEM_cv_fall))
         )

mmp <- ggplot(af3, aes(x = position, y = abs_diff, color = max_maf)) +
  geom_point() +
  scale_color_viridis_c()
mmp
```

And, now, let us also do a version in which the high max_mafs are printed on top, so that we can see them if they exist...
```{r}
mmp_sort <- ggplot(af3 %>% arrange(max_maf), 
                   aes(x = position, y = abs_diff, color = max_maf)) +
  geom_point() +
  scale_color_viridis_c()
mmp_sort
```

Outstanding.  It is clear that there is a fairly large haplotype/region
that is differentiated between fall and late fall there in the first
2 Mb of the chromosome.

## Display some of the highly associated SNPs there

The GWAS turned up a number of highly associated SNPs.  Let's add circles for some
of the ones with the largest likelihood ratio statistic.

We will take the top 6 from each of the comparisons:
```{r}
lfvsj <- read_tsv(input_list$gwas_lf_v_sanjo) %>%
  arrange(desc(LRT))
lfvfrhf <- read_tsv(input_list$gwas_lf_v_frhf) %>%
  arrange(desc(LRT))

# now get the top 6 from each for illustration
top6_from_each <- list(
  lfvsj %>% slice(1:6),
  lfvfrhf %>% slice(1:6)
) %>%
  bind_rows() %>%
  arrange(desc(LRT)) %>%
  filter(!duplicated(Position))

# print those out
top6_from_each
```

So, add those to the plot
```{r}
top6freqs <- af3 %>% semi_join(top6_from_each, by = c("position" = "Position"))
mmp_sort +
  geom_point(
    data = top6freqs,
    shape = 21,
    size = 3, 
    color = "black"
    )
```

That shows us that there are a lot of other markers with substantial allele freq
diffs in this region, but the association test puts thew big circles up as the most significant, perhaps
because of the sample sizes.  It sort of reinforces my informal belief that many insights
can come from looking at absolute allele freq diffs, at least in these sorts of
cases where you have a haplotypic region with
very large allele freq differences).

## Annotate the variants

Before we start designing assays and things, I want to look for large-effect mutations.
We have the annotations for the positions in these 5 Mb of Ots34 in the stored results
(from the Science paper), so we will use them here:

First we read in the annotations:
```{r}
ann_vcf <- read.vcfR(input_list$ots34_snpeff)

avt <- vcfR2tidy(ann_vcf, info_only = TRUE)

# get column heading for the snpEff output
ann_column_names <- avt$meta %>%
  filter(Tag == "INFO", ID == "ANN") %>%
  pull(Description) %>%
  str_replace_all(., pattern = "^.* '| +|'", "") %>%
  str_split(., "\\|") %>%
  pluck(1)

anns <- avt$fix %>%
  select(CHROM, POS, REF, ALT, ANN) %>%
  separate(ANN, into = ann_column_names, sep = "\\|")

```
I don't think we need to worry about the discarded pieces there.

Now, we can join those annotations onto the allele frequencies and plot them,
coloring by predicted effect:

```{r}
af3_ann <- left_join(af3, anns, by = c("position" = "POS")) %>%
  mutate(Annotation_Impact_f = factor(Annotation_Impact,
                                      levels = c("HIGH", "LOW", "MODERATE", "MODIFIER"))) %>%
  arrange(desc(Annotation_Impact_f))

imp_col <- ggplot(af3_ann, aes(x = position, y = abs_diff, colour = Annotation_Impact)) +
  geom_point() +
  scale_colour_manual(values = c(HIGH = "red", LOW = "green", MODERATE = "orange", MODIFIER = "gray90")) +
  theme_bw()
imp_col
```

So, there are a few that we would want to be sure to hit.

## Finding candidates for assay design

When I originally did this, I also used a hack to include indels in here.  I spare
the reader from the gory details of that.

We could rank things by allele frequency difference, but I would also like
to incorporate an estimate of the variance in these absolute frequency
differences.  The variance in that difference will be the sum of the
two variances, so it is easy to calculate.  We don't really know what the
sample size is, but I am going to throw out as an approximation that each
individual in the `nInd` field represents a sample of about 1.2 gene copies.

We will estimate the variances from each different estimate and then calculate
two standard deviations below the actual estimate and have that as a "lower
bound" on the absolute value of the allele frequency difference.

```{r}
# here is a function for computing the lower bound on the absolute alle freq diff
lub <- function(f1, f2, n1, n2, s = 1.2) {
  var <- (f1 * (1 - f1) / (n1 * s)) + (f2 * (1 - f2) / (n2 * s))
  sd <- sqrt(var)

  abs(f1 - f2) - 2 * sd
}

saind2 <- af3_ann %>%
  mutate(abs_diff_lb = lub(knownEM_late_fall, knownEM_cv_fall, nInd_late_fall, nInd_cv_fall))
```

Plot those to make sure they look right, and denote the most highly associated ones
so that we can see those. 
```{r}
top6again <- saind2 %>%
  semi_join(top6_from_each, by = c("position" = "Position"))

ggplot(saind2, aes(x = abs_diff, y = abs_diff_lb, colour = max_maf)) +
  geom_point() +
  scale_colour_viridis_c() +
  geom_point(data = top6again, fill = NA, colour = "black", size = 3, shape = 21)
```

All right.  When I look at that, I think to myself, The ones I would want to design
assays for (for distinguishing fall-run from late-fall), would be:

1. (1) The ones with the highest association values
2. (2) The ones with an abs_diff > 0.75 and an abs_diff "lower bound" above 0.5.  
3. (4) Then, for any remaining ones we might want to do, I would choose things with an
abs_diff > 0.5, and abs_diff_lb > 0.25, and a max_maf as close to 1 as possible.  
4. (8) Annotation_Impact %in% c("Moderate", "High")

We will designate those with a bit mask, with bits being 2 raised to their number 
minus 1 (as given in the parentheses, above).

**Another Category**  
I will also add another category of candidate, which is out in the flanks and might
be interesting for assessing LD out there.

* (16) Position > 2.2 Mb and abs_diff > 0.48 and abs_diff_lb > 0.25

Mark them.
```{r}
assay_candi <- saind2 %>%
  mutate(candi_category = 1 * position %in% top6_from_each$Position +
           2 * (abs_diff > 0.75 & abs_diff_lb > 0.5) +
           4 * (abs_diff > 0.5 & abs_diff_lb > 0.25 & max_maf > 0.98) + 
           8 * (Annotation_Impact %in% c("MODERATE", "HIGH") & (abs_diff > 0.5)) +
           16 * (abs_diff > 0.48 & abs_diff_lb > 0.20 & position > 2.2e6)) %>%
  select(chromo:max_maf, Annotation_Impact_f:candi_category, everything())

# how many?
assay_candi %>%
  count(candi_category)
```

Just to be explicit about what these candi_categories mean:

* 0 : not selected for any reason
* 1 : Hi-assoc (from Mac)
* 2 : abs_diff > 0.75 & abs_diff_lb > 0.5
* 3 : 1 and 2
* 4 : (abs_diff > 0.5 & abs_diff_lb > 0.25 & max_maf > 0.98)
* 7 : 1 and 2 and 4
* 8 : (Annotation_Impact %in% c("MODERATE", "HIGH") & (abs_diff > 0.5)) )
* 16 : abs_diff > 0.48 & abs_diff_lb > 0.20 & position > 2.2e6
* 20 : 4 and 16

Plot them on a plot like the above so we can see them.
```{r}
ggplot(mapping = aes(x = abs_diff, y = abs_diff_lb, colour = factor(candi_category))) +
  geom_point(data = assay_candi %>% filter(candi_category == 0), colour = "gray90") + 
  geom_point(data = assay_candi %>% filter(candi_category > 0)) +
  theme_bw() +
  scale_color_brewer(palette = "Dark2")

```

And see where those candidates live:
```{r}
ggplot(mapping = aes(x = position, y = abs_diff, colour = factor(candi_category))) +
  geom_point(data = assay_candi %>% filter(candi_category == 0), colour = "gray90") + 
  geom_point(data = assay_candi %>% filter(candi_category > 0)) +
  theme_bw() +
  scale_color_brewer(palette = "Dark2")
```

So, we have good candidates across a couple Mb.  That is great.

Let's make a plot of that for the supplement:
```{r}
assay_candi4fig <- assay_candi %>%
  mutate(
    candi_cat = case_match(
      candi_category,
      0 ~ NA_character_,
      1 ~ "1: One of the 10 SNPs with lowest p-values",
      2 ~ "2: |d| > 0.75 with lower confidence interval >0.5",
      3 ~ "3: 1 and 2",
      4 ~ "4: |d| > 0.5 with lower CI  >0.25 and near fixed in one group",
      7 ~ "7: 1 and 2 and 4",
      8 ~ "8: snpEff annotation of High or Moderate",
      16 ~ "16: |d| > 0.48 with lower CI  >0.2 and position >2.2 Mb",
      20 ~ "20: 4 and 16"
    ),
    `Candidate Marker Category` = factor(
      candi_cat,
      levels = c(
        "1: One of the 10 SNPs with lowest p-values",
        "2: |d| > 0.75 with lower confidence interval >0.5",
        "3: 1 and 2",
        "4: |d| > 0.5 with lower CI  >0.25 and near fixed in one group",
        "7: 1 and 2 and 4",
        "8: snpEff annotation of High or Moderate",
        "16: |d| > 0.48 with lower CI  >0.2 and position >2.2 Mb",
        "20: 4 and 16"
      ))
  )

g <- ggplot(mapping = aes(x = position, y = abs_diff, colour = `Candidate Marker Category`)) +
  geom_point(data = assay_candi4fig %>% filter(candi_category == 0), colour = "gray90") + 
  geom_point(data = assay_candi4fig %>% filter(candi_category > 0)) +
  theme_bw() +
  scale_color_brewer(palette = "Dark2") +
  xlab("Position on Chromosome 34 (Mb)") +
  ylab("|d|, Absolute Difference in Allele Frequency")

ggsave(g, filename = "tex/images/lfar-candidates.pdf", width = 8, height = 5)
```

## Preparing an assay order

We prepare assay order information for these 49 good candidates, and handed that
to Anthony, along with the sequences, to design primers.

That process involved some custom scripts and Primer3.  Once the candidates
came back, they were tested in a variety of Chinook samples and it was found that
only 8 gave reliable amplicons that mapped reliably, etc.  And those
are used in the ensuing work.

Here, print out the actual candidates:
```{r}
actual_candidates <- assay_candi %>%
  filter(candi_category > 0) %>% 
  arrange(position)

# write that out for fun:
write_csv(actual_candidates, file = output_list$candidates_for_assays)
```

# Citations

::: {#refs}
:::

# Session Info

```{r}
sessioninfo::session_info()
```

# Running Time

Running the code and rendering this notebook required approximately this much time:

```{r}
Sys.time() - start_time
```