### 2025-03-27 009-structure.Rmd
Done

---
title: "Running Structure"
author: "Eric C. Anderson"
date: "Last Updated: `r Sys.Date()`"
output:
  html_document:
    df_print: paged
    toc: true
    toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(message = FALSE)
start_time <- Sys.time()
```

# Introduction

Anthony ran these analyses by using the reference data set and running them
in the program Structure using the our lab's [slg_pipe](https://github.com/eriqande/slg_pipe)
framework for parallelization.

Subsequently, outputs were uploaded to the [CLUMPAK server](https://clumpak.tau.ac.il/) for
visualization.

# Session Info

```{r}
sessioninfo::session_info()
```

# Running Time

Running the code and rendering this notebook required approximately this much time:

```{r}
Sys.time() - start_time
```